### 2017-05-09 sheets/basis-2/sheet-basis-2.tex
Added some presheaf/contravariant excercises

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\DeclareMathOperator{\A}{\mathscr A}
\DeclareMathOperator{\B}{\mathscr B}
\DeclareMathOperator{\M}{\mathscr M}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 2}{Maps between Categories -- Functors}

\author{Mostafa Ahmed Abdelfattah Abouhamra, Felix Rech, Dominik Wagner}

\begin{hint}
  Read Chapter 1.2.
\end{hint}

\section{Contravariant Functors and Presheaves}

\begin{exercise}
  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories. Show that a contravariant functor $F:\A\rightarrow \B$ corresponds (uniquely) to a monotonically increasing function $f:A\rightarrow B$
\end{exercise}

\begin{exercise}
  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:

  \begin{enumerate}
  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
  \end{enumerate}

\end{exercise}

%\section{Diagrams}
%
%\begin{hint}
%  This is how you can draw diagrams:
%  \[
%    \begin{tikzcd}
%                                    & A \arrow{dr}{g}    & \\
%      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
%    \end{tikzcd}
%  \]
%\end{hint}

\section{Categories - Definition and Basics}

\begin {exercise}
Show there can be at most one inverse for a morphism $f \from A \to B$.
\end{exercise}

\begin {definition}{Rel}
The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
\[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
\end{definition}

\begin {exercise}
Show that \textbf{Rel} is a category.
\end{exercise}

\begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
(recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$). The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
\end{definition}

\begin {exercise}
Show that \textbf{Pos} is a category.
\end{exercise}

% \begin{exercise}
% For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
% \end{exercise}

\begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
\end{definition}

\begin{exercise}
  \begin{enumerate}
  \item Show that \textbf{Rel} is a pointed category.
  \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
  \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
  \end{enumerate}
\end{exercise}

\begin{exercise}
Show that any isomorphism is both monic and epic.
\end{exercise}

\newpage

\begin{exercise}
  Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
  \[
    \begin{tikzcd}
      A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
      & C &
    \end{tikzcd}
  \]
  Show the following facts:
  \begin{enumerate}
  \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
  \item If $h$ is monic, then $f$ is monic.
  \item If $h$ is epic, then $g$ is epic.
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
  \begin{enumerate}
  \item If $f$ is monic, then it is injective.
  \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
  \end{enumerate}
\end{exercise}

\begin{exercise}[Challenge 2]
  Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
  \begin{hint}
    Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
  \end{hint}
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: