### 2017-05-10 sheets/basis-2/sheet-basis-2.tex
Added extended version of exercises.

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\DeclareMathOperator{\A}{\mathscr A}
\DeclareMathOperator{\B}{\mathscr B}
\DeclareMathOperator{\M}{\mathscr M}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 2}{Maps between Categories -- Functors}

\author{Mostafa Ahmed Abdelfattah Abouhamra, Felix Rech, Dominik Wagner}

\begin{hint}
  Read Chapter 1.2.
\end{hint}

\section{Maps between Categories -- Functors}

\begin{exercise}
  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories. Show that a contravariant functor $F:\A\rightarrow \B$ corresponds (uniquely) to a monotonically increasing function $f:A\rightarrow B$
\end{exercise}

\begin{exercise}
  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:

  \begin{enumerate}
  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Not entirely trivial 1)}
  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
  \end{enumerate}

\end{exercise}

\begin{definition}{Coslice Category}
  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{coslice category} $A/\A$ of $\A$ under $A$ consists of
  \begin{itemize}
  \item the arrows to $A$ (i.e. the codomain of which are $A$) as objects
  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
  \end{itemize}
\end{definition}

\begin{definition}{$\mathbf{Sets_*}$}
  The category $\mathbf{Sets_*}$ of \emph{pointed sets} consists of
  \begin{itemize}
  \item pairs $(A,a)$ such that $A$ is a set and $a\in A$ as objects and
  \item arrows objects between objects $(A,a)$ and $(B,b)$ are functions $f\from A\to B$ such that $f(a)=b$.
  \end{itemize}
\end{definition}

\begin{exercise}
Show that $\mathbf{Sets_*\cong 1/\mathbf{Sets}}$ where $1$ is an arbitrary set with exactly one element.  
\end{exercise}

\begin{definition}{Slice Category}
  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{slice category} $\A/A$ of $\A$ over $A$ consists of
  \begin{itemize}
  \item the arrows from $A$ (i.e. the domain of which are $A$) as objects
  \item and for $f\in \A(A,B)$, $g\in\A(A',B)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
  \end{itemize}
\end{definition}

\begin{exercise}[\textit{Challenge}]
  Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this indeed a functor. Is this also an isomorphism?

  \begin{hint}
    For arrows $f\in\A(A,B)$ consider functors $F_f\from\A/A\to\A/B$ that act on objects $g\in \A(C,A)$ by $F_f(g)=f\of g$. 
  \end{hint}
\end{exercise}

% \begin{exercise}[\textit{Challenge}]
% Let $\A$ be a category. Consider again the functor $F_B\from \A\to\mathbf{Hom}(-,B)$ that is defined for a fixed $B\in\ob(A)$ by
% \begin{align*}
%   F_B(A)&=\A(A,B)\\
%   F_B(f)&=F_{B,f}
% \end{align*}
% for all $A,A'\in\ob(\A)$, $f\in \A(A,A')$ and
% \begin{align*}
%   F_{B,f}(g)=g\of f
% \end{align*}
% for all $g\in\A(A,B)$.

% Fill in the details of the construction of the functor $F\from \A^{\op} \to \mathbf{CAT}$ that satisfies $F(B)=\mathbf{Hom}(-,B)$, i.e. define its behaviour on arrows and verify that this is indeed a functor.

% \begin{hint}
%   Consider the functors $F_f\from\mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ for $f\in\A(B',B)$ given by
%   \begin{align*}
%     F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
%     F_f(g)&=F_{f,g}
%   \end{align*}
%   where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and the function $F_{f,g}$ is defined by
%   \begin{align*}
%     F_{f,g}(h)= g(f\of h).
%   \end{align*}
%   for each arrow $h\from\A(A,B')$.
% \end{hint}

% % Is this functor also an isomorphism

% % given by
% %   \begin{align*}
% %     F(B)&=\mathbf{Hom}(-,B)\\
% %     F(f)&=F_f
% %   \end{align*}
% % where $A\in\ob(\A)$ and $f\in\A(B',B)$, each functor $F_f\from \mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ is given by
% % \begin{align*}
% %   F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
% %   F_f(g)&=F_{f,g}
% % \end{align*}
% % where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and
% % \begin{align*}
% %   F_{f,g}(h)= g(f\of h).
% % \end{align*}
% % for each function $h\from\A(A,B)\to\A(A',B)$
% % Check that this is indeed a functor, i.e. satisfies the axioms, and determine whether it is also an isomorphism.
% \end{exercise}

% %\section{Diagrams}
% %
% %\begin{hint}
% %  This is how you can draw diagrams:
% %  \[
% %    \begin{tikzcd}
% %                                    & A \arrow{dr}{g}    & \\
% %      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
% %    \end{tikzcd}
% %  \]
% %\end{hint}

% \section{Categories - Definition and Basics}

% \begin {exercise}
% Show there can be at most one inverse for a morphism $f \from A \to B$.
% \end{exercise}

% \begin {definition}{Rel}
% The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
% The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
% Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
% \[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
% \end{definition}

% \begin {exercise}
% Show that \textbf{Rel} is a category.
% \end{exercise}

% \begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
% (recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$). The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
% \end{definition}

% \begin {exercise}
% Show that \textbf{Pos} is a category.
% \end{exercise}

% % \begin{exercise}
% % For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
% % \end{exercise}

% \begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
% \end{definition}

% \begin{exercise}
%   \begin{enumerate}
%   \item Show that \textbf{Rel} is a pointed category.
%   \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
%   \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}
% Show that any isomorphism is both monic and epic.
% \end{exercise}

% \newpage

% \begin{exercise}
%   Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
%   \[
%     \begin{tikzcd}
%       A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
%       & C &
%     \end{tikzcd}
%   \]
%   Show the following facts:
%   \begin{enumerate}
%   \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
%   \item If $h$ is monic, then $f$ is monic.
%   \item If $h$ is epic, then $g$ is epic.
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}
%   Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
%   \begin{enumerate}
%   \item If $f$ is monic, then it is injective.
%   \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}[Challenge 2]
%   Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
%   \begin{hint}
%     Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
%   \end{hint}
% \end{exercise}

\end{document}

% %%% Local Variables:
% %%% mode: latex
% %%% TeX-master: t
% %%% End: