### 2017-05-19 sheets/basis-3/sheet-basis-3.tex
Fixed the diagram of 3.4 to match the orientation of n.t. diagrams in Leinster.

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}

\begin{hint}
  Read Chapter 1.3.
\end{hint}

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}

\begin{exercise}
        Let $\cat{A}$ and $\cat{B}$ be categories.  Prove that
        $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
\end{exercise}

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like 'God-given' or
  'defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}

\begin{definition}{Product of Morphisms inside a Category}
  Let $\cat{A}$ be a category and $A, B, C, D \in \ob(\cat{A})$ with $f \from A \to C$ and $g \from B \to D$.
  The product morphism of $f$ and $g$, written $f \times g$, is a function
  \begin{align*}
    A \times B &\qquad\to\qquad C \times D\\
    (a, b) &\qquad\mapsto\qquad (f(a), g(b))
  \end{align*}
\end{definition}

\newcommand{\eval}[1]{\ensuremath{\mathsf{eval}^{#1}\xspace}}

\begin{exercise}
  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
  a functor $F_A:\mathbf{Set} \to \mathbf{Set}$ as follows:
  \begin{align*}
    F_A(B) & =  B^A \times A\\
    F_A(f: B \to C) & =  (f \circ \_) \times 1_A
  \end{align*}
  The function $\eval{A} \from F_A \to 1_{\mathbf{Set}}$, defined as $\eval{A}(f,a) := f(a)$, is a natural transformation if the following diagram commutes for each $g \from B \to C$.
\[
  \begin{tikzcd}
    F_A(B) = B^A \times A \arrow{rrr}{F_A(g)=(g\circ \_) \times 1_A} \arrow{dd}[swap]{\eval{A}_B} & & & 1_{\mathbf{Set}}(B) = B \arrow{dd}{\eval{A}_C} \\
                                                                                           & & & \\
    F_A(C) = C^A \times A \arrow{rrr}[swap]{1_{\mathbf{Set}}(g) = g} & & & 1_{\mathbf{Set}}(C) = C
  \end{tikzcd}
\]

Check that this diagram commutes.
\end{exercise}

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
                \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
                \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}

\begin{exercise}
A \emph{permutation} of a set $X$ is a bijection $X \to X$.
Write $\Sym(X)$ for the set of permutations of $X$.
A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
either $x \leq y$ or $y \leq x$.
Write $\Ord(X)$ for the set of total orders on $X$.

Let $\cat{B}$ denote the category of finite sets and bijections.

        \begin{itemize}
        \item[(a)]
Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
definitions should be canonical (no arbitrary choices).

        \item[(b)]
Show that there is no natural transformation $\Sym \to \Ord$.
(Hint: consider identity permutations.)

        \item[(c)]
For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
and $\Ord(X)$ have?
        \end{itemize}

Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
set $X$, there are exactly as many permutations of $X$ as there are total
orders on $X$, but there is no natural way of matching them up.)
\end{exercise}

\begin{exercise}
Let $F\from \cat{A} \to \cat{B}$ be a functor.

        \begin{itemize}
        \item[(a)]
Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
essentially surjective on objects.  (Hint: prove faithfulness before
fullness.)

        \item[(b)]
Now suppose instead that $F$ is full, faithful and essentially surjective
on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
and an isomorphism $\varepsilon\from F(G(B)) \to B$.  Prove that $G$ extends
to a functor in such a way that $(\varepsilon)_{B \in \cat{B}}$ is a natural
isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
        \end{itemize}
\end{exercise}

\begin{exercise}
This exercise makes precise the idea that linear algebra can equivalently
be done with matrices or with linear maps.

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: