### 2017-05-19 sheets/basis-2/sheet-basis-2.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\newcommand{\A}{\cat A}
\newcommand{\B}{\cat B}
\newcommand{\M}{\cat M}
\newcommand{\Rel}{\fcat{Rel}}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 2}{Maps between Categories -- Functors}

\author{Mostafa Ahmed Abdelfattah Abouhamra, Felix Rech, Dominik Wagner}

\begin{hint}
  Read Chapter 1.2.
\end{hint}

\begin{exercise}
Show that functors preserve isomorphism. That is, prove that if $F : \cat{A} \to \cat{B}$ is a functor and $A, A' \in \cat{A}$ with $A \iso A'$, then $F(A) \iso F(A')$.
\end{exercise}

\begin{answer}
  Since $A\cong A'$ there are $f\in\mathscr A(A,A')$ and $g\in\mathscr A(A',A)$ such that $g \of f=1_A$ and $f \of g=1_A'$. Hence, $F(g)\of F(f)=F(g\of f)=F(1_A)=1_{F(A)}$ and $F(f)\of F(g)=F(f\of g)=F(1_A')=1_{F(A')}$. As a consequence, $F(f)$ is an isomorphism between $F(A)$ and $F(A')$. In particular, $F(A)\cong F(A')$.
\end{answer}

\begin{exercise}
  Check that there is a category $\Cat$ with categories as objects and functors as arrows.
\end{exercise}

\begin{answer}
In $\mathbf{CAT}$, objects are categories and the self loops are trivial endofunctors and so they trivially follow identity laws.\\
This counts as a trivial answer to the question. However, we can also include all functors.\\
For associativity, consider three functors $F:A\rightarrow B$, $G:B\rightarrow C$ and $H:C\rightarrow D$.\\
\begin{itemize}
\item $(H \of G) \of F(a)$ = $H \of G(F(a))$ = $H(G(F(a)))$
\item $H \of (G \of F) (a)$ : $H \of (G \of F(a))$ = $G \of G(F(a))$ = H(G(F(a)))
\item Therefore, $(H \of G) \of F(a) = H \of (G \of F) (a)$ and we satisfy associativity for objects.
\item The same works for morphisms because $F(f' \of f)  = F(f') \of F(f)$ by definition. Therefore, we satisfy associativity for objects.
\item Since identities are mapped to identities, we satisfy identity laws
\item Therefore, $\mathbf{CAT}$ is a category
\end{itemize}
\end{answer}

\begin{exercise}
  Find an example of a functor $F : \cat{A} \to \cat{B}$ such that $F$ is faithful but there exist distinct maps $f_1$ and $f_2$ in $\cat{A}$ with $F(f_1) = F(f_2)$.
\end{exercise}

\begin{answer}
  Suppose the discrete categories $\A$ and $\B$ with two and one objects respectively. Then, any functor $F\from A\to B$ is trivially faithful but $F(1_A)=F(1_B)$ (where $A$ and $B$ are the only objects of $\A$).
\end{answer}

\begin{exercise}
  Find interesting examples for a functor that is both full and faithful, one that is full but not faithful, one that is faithful but not full, and one that is neither.
\end{exercise}

\begin{answer}
\begin{itemize}
\item Any bijective functor from a discrete category to another is a full and faithful functor.
\item A functor that maps a category with two objects and two morphisms (both have the same domain and codomain) to a category with the same two objects and one morphism is full, but not faithful.
\item A functor that maps a category with two objects and one morphism to a category with the same two objects and two morphisms is faithful, but not full.
\item Combining the last two to have a category with four objects which are divided into two pairs as mentioned before which are mapped to another category with four objects that are divided into two pairs with the same mentioned structures.
\end{itemize}
\end{answer}

\begin{exercise}
  \begin{enumerate}
    \item What are the subcategories of an ordered set? Which are full?
    \item What are the subcategories of a group? (Careful!)
      Which are full?
  \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item A subcategory of an ordered set $(S,\leq)$ is an ordered set $(S',\leq')$ such that $S'\subseteq S$ and $\leq'\;\subseteq\;\leq$. In the case of full subcategories even for all $s,t\in S'$, $s\leq t$ if and only if $s\leq' t$ holds.
  \item The subcategories of a group are either the category with no objects or submonoids of the group (which is of course a monoid). The only full subcategories are the group itself and the category with no objects.
  \end{enumerate}
\end{answer}

\begin{exercise}
  \begin{enumerate}
    \item Define an endofunctor on $\Set$ that maps every set to its powerset.
      Check that it satisfies the functor laws.
      Is it full/faithful?
    \item Define a contravariant endofunctor on $\Set$, i.e.\ a functor from $\Set^\op$ to $\Set$, that maps every set to its powerset.
      Check that it satisfies the functor laws.
      Is it full/faithful?
  \end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item
\begin{itemize}
\item Since every set is mapped to another set (its powerset), then all objects are mapped.
\item A function $f: X \rightarrow Y$ is mapped to $F: P(X) \rightarrow P(Y)$, such that $F(x)  = \{x \in X\}$.
\item This makes the axioms follow easily from composition in $\mathbf {Set}$. Therefore, it satisfies functor laws.
\item It is faithful (A set has only one powerset) and not full (consider a function that takes the entire powerset as an element and maps it to the other powerset).
\end{itemize}
\item
The only difference is mapping elements in the powerset to their preimage set. The rest of the argument is the same.
\end{enumerate}
\end{answer}

\begin{exercise}
  Let $A$ and $B$ be posets, and $\A$ and $\B$ be the corresponding categories. Consider a contravariant functor $F:\A^{\op}\rightarrow \B$. Describe properties of the corresponding function $f:A\rightarrow B$.
\end{exercise}

\begin{answer}
  For every arrow $x \to y$ in $\A$, there is an arrow $f : y \to x$ in $\A^\op$ and hence an arrow $F(f) : F(y) \to F(x)$ in $\B$.
  In the language of posets this means that $x \leq y$ implies $f(x) \geq f(y)$, i.e.\ $f$ is monotonically decreasing.
  The functor laws hold tautologically in posets and don't tell us anything about $f$.
\end{answer}

\begin{exercise}
  Determine which of the following holds, i.e.\ either provide an isomorphism in $\mathbf{CAT}$ or argue why there is none:

  \begin{enumerate}
  \item $\mathbf {Rel}^{\op}\cong \mathbf {Rel}$;
  \item $\mathbf {Set}^{\op}\cong \mathbf {Set}$;
  \item for an arbitrary monoid interpreted as a category $\M$ with one object:  $\M^{\op}\cong \M$; \textit{(Challenge 1)}
  \item for $\mathbf{CAT_i}$, the subcategory of $\mathbf{CAT}$ consisting of all categories and functors that are isomorphisms: $\mathbf{CAT_i}^{op}\cong\mathbf{CAT_i}$.
  \end{enumerate}

\end{exercise}

\begin{answer}
  \begin{enumerate}
    \item The categories are isomorphic.
      We define functors $F : \Rel^\op \to \Rel$ and $G : \Rel \to \Rel^\op$ in both directions by
      \begin{align*}
        F(X) &\coloneqq X  &  G(X) &\coloneqq X \\
        F(R) &\coloneqq R^{-1}  &  G(R) &\coloneqq R^{-1}
      \end{align*}
      for all sets $X$ and $Y$ and relations $R \in X \times Y$.
      Both functors are mutually inverse bacause $(R^{-1})^{-1} = R$.
      Thus $\Rel^\op \iso \Rel$.

    \item We refute the statement by contradiction.
      Assume that we have an isomorphism $F : \Set^\op \to \Set$.
      Let $1$ be a one-element set.
      It is final in $\Set$ and hence initial in $\Set^\op$ and will be mapped to the initial object $\emptyset$ in $\Set$, because isomorphisms preserve initial objects.
      We define $f : 1 \to X$ as a function to some set $X$ that is not isomorphic to $1$.
      Then $F(X)$ is not isomorphic to $F(1) = \emptyset$, but there is a function $F(f) : F(X) \to \emptyset$, which leads to a contradiction.

    \item Let $X$ be a finite set with at least two elements.
      We define the monoid $\M \coloneqq \lbrace f : X \to X \rbrace$ with function composition as binary operation and the identity function as neutral element.
      Let $c \in \M$ be some constant function.
      Such a function is left-absorbing, which means that for all functions $f \in \M$, we have $c \circ f = c$.
      Left-absorbing elements in the opposite monoid correspond to right-absorbing functions, i.e., fuctions $d \in \M$ such that for all $f \in M$, we have $f \circ d = d$.
      Obviously such functions do not exist.
      Since left-absorbing elements are preserved by isomorphisms, $\M$ cannot be isomorphic to its opposite monoid.

    \item We define mutually inverse functors $F : \CAT_i^\op \to \CAT_i$ and $G : \CAT_i \to \CAT_i^\op$ by
      \begin{align*}
        F(\A) &\coloneqq \A  &  G(\A) &\coloneqq \A \\
        F(H) &\coloneqq H^{-1}  &  G(H) &\coloneqq H^{-1}
      \end{align*}
      for all categories $\A$ and $\B$ and functors $H : \A \to B$.
      The functors $F$ and $G$ are mutually inverse because $(H^{-1})^{-1} = H$.
      Thus $\CAT_i^\op \iso \CAT_i$.
  \end{enumerate}
\end{answer}

\begin{definition}{Free Monoid}
  For every set $A$, we define the \emph{free monoid} $A^* \coloneqq \lbrace \text{words over $A$} \rbrace$.
  We define the binary operation by $w \cdot w' \coloneqq ww'$ as the concatenation of words and use the empty word $\varepsilon$ as neutral element.
\end{definition}

\begin{exercise}
  Show that for every set $A$, the free monoid $A^*$ satisfies the following \emph{universal mapping property}:
  There is a function $i : A \to A^*$ and given any monoid $M$ and any function $j : A \to M$, there is a \emph{unique} monoid homomorphism $f : A^* \to M$ such that $f \circ i = j$.
\end{exercise}

\begin{answer}
  Define
  \begin{align*}
    i(a)&=a &(a\in A)
  \end{align*}
  and
  \begin{align*}
    f(\epsilon)&=0_M&\\
    f(a_1\cdots a_{n-1}a_n)&=f(a_1\cdots a_{n-1})\of j(a_n) &\text{for }n>0.
  \end{align*}
  Clearly, for each $a\in A$, $f(i(a))=j(a)$, thus $f\of i=j$, and $f$ is a group homomorphism.

  Now, suppose $g$ is a group homomorphism satisfying $g\of i=j$.
  Let $a_1\cdots a_n\in A^*$. Proof by induction on $n$ that $g(a_1\cdots a_n)=f(a_1\cdots a_n)$.

  Base case: $g(\epsilon)=0_M=f(\epsilon)$ because $\epsilon$ is the neutral element in $A^*$).

  Inductive step:
  \begin{align*}
    g(a_1\cdots a_{n-1}a_n)&=g(a_1\cdots a_{n-1})\of g(a_n)&g\text{ group homomorphism}\\
                           &=f(a_1\cdots a_{n-1})\of g(a_n)&\text{inductive hypothesis}\\
                           &=f(a_1\cdots a_{n-1})\of g(i(a_n)) & a_n\in A\\
                           &=f(a_1\cdots a_{n-1})\of f(i(a_n)) & g\of i=j=f\of i\\
                           &=f(a_1\cdots a_{n-1})\of f(a_n) & a_n\in A\\
                           &=f(a_1\cdots a_{n-1}a_n)&f\text{ group homomorphism}
  \end{align*}
  Hence, $f=g$, so $f$ is indeed unique.

\end{answer}

\begin{exercise}
  Denote by $\Mon$ the category of monoids with monoid morphisms as arrows.
  Define a functor $F : \Set \to \Mon$ that maps every set $A$ to the free monoid $A^*$.
  Is $F$ full/faithful?
\end{exercise}

\begin{answer}
\begin{itemize}
\item We assume mapping a function $f: X \rightarrow Y in$ $\mathbf {Set}$ to a free monoid morphism $F: MonX \rightarrow MonY$ such that $F(Z \in MonX) = (f(z) | z \in Z)$
\item F is faithful since if two sets have the same free monoids (All the lists that can be generated from the elements of those sets), then the two sets are the same.
\item F is not full because we can have a morphism $F(Z \in MonX) = (f(z)f(z) | z \in Z)$
\end{itemize}
\end{answer}

\begin{exercise}
  \begin{enumerate}
    \item A directed graph consists of a set of vertices $V$, a set of edges $E$ and two functions $s, t : E \to V$ that assign a source and target vertex to each edge.
      Define a category $\cat{A}$ such that functors from $\cat{A}$ to $\Set$ correspond to directed graphs.
    \item Define a category $\cat{B}$ such that the functors from $\cat{B}$ to $\Set$ correspond to injections between an arbitrary pair of sets.
  \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item Consider $\A$ given by
    \[
      \begin{tikzcd}[column sep=huge,row sep=huge]
        E\arrow[r,yshift=.85ex,"s"] \arrow[r,below,yshift=-.85ex,"e"]&V
      \end{tikzcd}
    \]
    Then a functor $F\from \A\to\Set$ assigns to both $E$ and $V$ sets $E'$ and $V'$ (which correspond to the set of edges and vertices respectively) and to $s,e$ functions $s'$ and $e'$ that take edges $d$ from $E'$ and return elements $s'(d)$ and $e'(d)$ which correspond to the start and end node of the edge $d$ respectively.
  \item Consider $\B$ given by
    \[
      \begin{tikzcd}[column sep=huge,row sep=huge]
        A\arrow[r,yshift=.85ex,"f"] & B\arrow[l,below,yshift=-.85ex,"g"] \arrow[loop right,"h"]
      \end{tikzcd}
    \]
    (plus the identities) such that $g\of f=1_A$, $f\of g=h$, $g\of h=g$ and $h\of f=f$. Then any functor $F\from\A\to\Set$ maps $A$ and $B$ to sets $A'$ and $B'$ respectively and $f$ is mapped to a function $f'\from A'\to B'$ and $g$ is mapped to a function $g'\from B'\to A'$ such that for all $a\in A'$, $g'(f'(a))=a$. Thus $f'$ is an injection.

    Indeed, the reverse is also true: Given any injection $f'\from A'\to B'$ there is functor $F\from \A\to\Set$ such that $f'$ corresponds to $F$ (in the way specified above).
  \end{enumerate}
\end{answer}

\begin{definition}{Coslice Category}
  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{coslice category} $A/\A$ of $\A$ under $A$ consists of
  \begin{itemize}
  \item the arrows from $A$ (i.e. the domain of which is $A$) as objects
  \item and for $f\in \A(A,B)$, $g\in\A(A,B')$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $h\of f=g$.
  \end{itemize}
\end{definition}

\begin{definition}{$\mathbf{Sets_*}$}
  The category $\mathbf{Sets_*}$ of \emph{pointed sets} consists of
  \begin{itemize}
  \item pairs $(A,a)$ such that $A$ is a set and $a\in A$ as objects and
  \item arrows between objects $(A,a)$ and $(B,b)$ are functions $f\from A\to B$ such that $f(a)=b$.
  \end{itemize}
\end{definition}

\begin{exercise}
Show that $\mathbf{Sets_*\cong 1/\mathbf{Sets}}$ where $1$ is an arbitrary set with exactly one element.
\end{exercise}

\begin{answer}
  We define a functor $F : \Set_* \to 1/\Set$ by
  \begin{align*}
    F((A, a)) &\coloneqq (x \mapsto a) \\
    F(h) &\coloneqq h
  \end{align*}
  for all pointed sets $(A, a)$ and $(B, b)$ and functions $h : A \to B$ with $f(a) = b$.
  The second line is valid because
  \[ \begin{split}
    h \circ F((A, a))
    &= (x \mapsto h(a)) \\
    &= (x \mapsto b) \\
    &= F((B, b)).
  \end{split} \]
  Now we define the inverse functor $G : 1/\Set \to \Set_*$ by
  \begin{align*}
    G(f) &\coloneqq (A, f(\star)) \\
    G(h) &\coloneqq h
  \end{align*}
  for all functions $f : 1 \to A$, $g : 1 \to B$ and $h : A \to B$ with $h \circ f = g$.
  The second line is valid because
  \[ \begin{split}
    h(f(\star))
    &= (h \circ f)(\star) \\
    &= g(\star).
  \end{split} \]
  Both functors are mutually inverse on objects because
  \[ \begin{split}
    F(G(f))
    &= F((A, f(\star))) \\
    &= (x \mapsto f(\star)) \\
    &= f
  \end{split} \]
  and
  \[ \begin{split}
    G(F((A, a)))
    &= G(x \mapsto a) \\
    &= (A, a).
  \end{split} \]
  On arrows they are trivially inverse.
\end{answer}

\begin{definition}{Slice Category}
  Let $\A$ be a category and $A\in\ob(\A)$ an object. The \emph{slice category} $\A/A$ of $\A$ over $A$ consists of
  \begin{itemize}
  \item the arrows to $A$ (i.e. the codomain of which is $A$) as objects
  \item and for $f\in \A(B,A)$, $g\in\A(B',A)$ (which are objects in $A/\A$) the arrows between $f$ and $g$ in $A/\A$ are arrows $h$ such that $g\of h=f$.
  \end{itemize}
\end{definition}

\begin{exercise}[\textit{Challenge 2}]
  Let $\A$ be a category. Construct a functor $F\from \A\to\mathbf{Cat}$ such that $F(A)=\A/A$ and show that this is indeed a functor. Is this also an isomorphism?

  \begin{hint}
    For arrows $f\in\A(A,B)$ consider functors $F_f\from\A/A\to\A/B$ that act on objects $g\in \A(C,A)$ by $F_f(g)=f\of g$.
  \end{hint}
\end{exercise}

\begin{answer}
  As suggested, we define
  \begin{align*}
    F(A) &\coloneqq \A / A \\
    F(f) &\coloneqq F_f,
  \end{align*}
  where $F_f : \A / A \to \A / B$ is given by
  \begin{align*}
    F_f(g) &\coloneqq f \circ g \\
    F_f(h) &\coloneqq h
  \end{align*}
  for all functions $g$, $g'$ and $h$ such that the following diagram commutes.
  \[ \begin{tikzcd}[column sep=tiny]
    C \arrow[dr, "g"'] && C' \arrow[ll, "h"] \arrow[dl, "g'"] \\
    & A \arrow[d, "f"] \\
    & B
  \end{tikzcd} \]
  The second line is valid because
  \[ \begin{split}
    F_f(g) \circ h
    &= f \circ g \circ h \\
    &= f \circ g' \\
    &= F_f(g')
  \end{split} \]
  Our definition of $F$ satisfies both functor laws:
  \begin{enumerate}
    \item For $f : A_2 \to A_3$ and $f' : A_1 \to A_2$, we have $F(f \circ f') = F(f) \circ F(f')$:
      \begin{align*}
        F_{f \circ f'}(g)
          &= f \circ f' \circ g \\
          &= F_f(F_f'(g)) \\
        F_{f \circ f'}(h)
          &= h \\
          &= F_f(F_f'(h))
      \end{align*}
    \item We have $F(1_A) = 1_{F(A)}$:
      \begin{align*}
        F_{1_A}(g) &= g \\
        F_{1_A}(h) &= h
      \end{align*}
  \end{enumerate}
  In general the functor $F$ is not an isomorphism.
  For example, if $\A$ is the empty category then it is not isomorphic to $\Cat$.
\end{answer}

\end{document}

% \begin{exercise}[\textit{Challenge}]
% Let $\A$ be a category. Consider again the functor $F_B\from \A\to\mathbf{Hom}(-,B)$ that is defined for a fixed $B\in\ob(A)$ by
% \begin{align*}
%   F_B(A)&=\A(A,B)\\
%   F_B(f)&=F_{B,f}
% \end{align*}
% for all $A,A'\in\ob(\A)$, $f\in \A(A,A')$ and
% \begin{align*}
%   F_{B,f}(g)=g\of f
% \end{align*}
% for all $g\in\A(A,B)$.

% Fill in the details of the construction of the functor $F\from \A^{\op} \to \mathbf{CAT}$ that satisfies $F(B)=\mathbf{Hom}(-,B)$, i.e. define its behaviour on arrows and verify that this is indeed a functor.

% \begin{hint}
%   Consider the functors $F_f\from\mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ for $f\in\A(B',B)$ given by
%   \begin{align*}
%     F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
%     F_f(g)&=F_{f,g}
%   \end{align*}
%   where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and the function $F_{f,g}$ is defined by
%   \begin{align*}
%     F_{f,g}(h)= g(f\of h).
%   \end{align*}
%   for each arrow $h\from\A(A,B')$.
% \end{hint}

% % Is this functor also an isomorphism

% % given by
% %   \begin{align*}
% %     F(B)&=\mathbf{Hom}(-,B)\\
% %     F(f)&=F_f
% %   \end{align*}
% % where $A\in\ob(\A)$ and $f\in\A(B',B)$, each functor $F_f\from \mathbf{Hom}(-,B)\to\mathbf{Hom}(-,B')$ is given by
% % \begin{align*}
% %   F_f(\A(A,B))&=\{f\of g\mid g\in \A(A,B)\}\\
% %   F_f(g)&=F_{f,g}
% % \end{align*}
% % where $g\in\mathbf{Hom}(-,B)(\A(A,B),\A(A',B))$ and
% % \begin{align*}
% %   F_{f,g}(h)= g(f\of h).
% % \end{align*}
% % for each function $h\from\A(A,B)\to\A(A',B)$
% % Check that this is indeed a functor, i.e. satisfies the axioms, and determine whether it is also an isomorphism.
% \end{exercise}

% %\section{Diagrams}
% %
% %\begin{hint}
% %  This is how you can draw diagrams:
% %  \[
% %    \begin{tikzcd}
% %                                    & A \arrow{dr}{g}    & \\
% %      B \arrow{ur}{f} \arrow{rr}{h} &                    & C
% %    \end{tikzcd}
% %  \]
% %\end{hint}

% \section{Categories - Definition and Basics}

% \begin {exercise}
% Show there can be at most one inverse for a morphism $f \from A \to B$.
% \end{exercise}

% \begin {definition}{Rel}
% The objects of the category \textbf{Rel} are sets. The morphisms $f \from A \to B$ are subsets $f \sub A \times B$.
% The identity morphism on set $A$ is the equality relation $\{\round{a,a} \such a \in A\} := 1_A$.
% Composition of two morphisms $f \sub A \times B$, $g \sub B \times C$ is defined as
% \[ g \of f = \{ \round{a, c} \in A \times C \such \exists b. \round{a, b} \in f \land \round{b, c} \in g\} \]
% \end{definition}

% \begin {exercise}
% Show that \textbf{Rel} is a category.
% \end{exercise}

% \begin{definition}{Pos} The objects of  \textbf{Pos} are partially ordered sets
% (recall that a poset is a set $A$ equipped with a reflexive, transitive and antisymmetric binary relation $\leq^A$). The morphisms $m \from A \to B $ are monotone functions: $a \leq^A a' \implies m(a) \leq^A m(a')$
% \end{definition}

% \begin {exercise}
% Show that \textbf{Pos} is a category.
% \end{exercise}

% % \begin{exercise}
% % For a fixed set X with Powerset $\mathcal{P}(X)$, does $\mathcal{P}(X) \cong \mathcal{P}(X)^{op}$ hold?
% % \end{exercise}

% \begin{definition}{Pointed Category} If an object is both \emph{initial} and \emph{terminal}, it is called a \emph{zero object}. A \emph{pointed category} is one with a zero object.
% \end{definition}

% \begin{exercise}
%   \begin{enumerate}
%   \item Show that \textbf{Rel} is a pointed category.
%   \item Show that the category \textbf{Grp} of groups has both an initial and a final object, and that these are the same.
%   \item Show that the category \textbf{Ring} of unital rings has both an initial and a final object, and that these are \emph{not} the same.
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}
% Show that any isomorphism is both monic and epic.
% \end{exercise}

% \newpage

% \begin{exercise}
%   Let $A, B, C$ be objects in a category and let $f \from A \to B$ and $g \from B \to C$ be morphisms such that the following diagram commutes:
%   \[
%     \begin{tikzcd}
%       A \arrow{r}{f} \arrow{dr}{h} & B \arrow{d}{g} & \\
%       & C &
%     \end{tikzcd}
%   \]
%   Show the following facts:
%   \begin{enumerate}
%   \item If both $f$ and $g$ are isomorphisms, then $h$ is also an isomorphism.
%   \item If $h$ is monic, then $f$ is monic.
%   \item If $h$ is epic, then $g$ is epic.
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}
%   Show the following for a morphism $f \from A \to B$ in the category \textbf{Set}:
%   \begin{enumerate}
%   \item If $f$ is monic, then it is injective.
%   \item If $f$ is epic, then it is surjective. \textit{(Challenge 1)}
%   \end{enumerate}
% \end{exercise}

% \begin{exercise}[Challenge 2]
%   Show that in the category \textbf{Ring}, there are morphisms which are epic but not surjective.
%   \begin{hint}
%     Consider the inclusion morphism $\mathbb{Z} \to \mathbb{Q}$, and use the properties of ring homomorphisms.
%   \end{hint}
% \end{exercise}

\end{document}

% %%% Local Variables:
% %%% mode: latex
% %%% TeX-master: t
% %%% End: