### 2017-05-22 sheets/basis-3/sheet-basis-3.tex
Some solutions

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}

\begin{hint}
  Read Chapter 1.3.
\end{hint}

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}
\begin{answer}
    \begin{minipage}{0.5\textwidth}
        $\cat{A}$:
        \[
            \begin{tikzcd}
                A \arrow{d}{f}\\
                B
            \end{tikzcd}
        \]
    \end{minipage}
    \begin{minipage}{0.5\textwidth}
        $\cat{B}$:
        \[
            \begin{tikzcd}
                A' \arrow{dr}{f'} \arrow[bend left]{rr}{g} &    & A'' \arrow{dl}{f''} \arrow{ll}{g^{-1}} \\
                                                           & B' &
            \end{tikzcd}
        \]
    \end{minipage}
    The identity morphisms have been left out for simplicity.
    First of all note that there is no bijection between the objects of $\cat{A}$ and $\cat{B}$, so $\cat{A} \not\iso \cat{B}$.\\
    We now show $\cat{A} \eqv \cat{B}$.
    Therefore we define a functor $F$:
    \begin{align*}
        F(A) := A'\\
        F(B) := B'\\
        F(f) := f'
    \end{align*}
    In addition $F$ maps the identity morhisms to their respective identity morphisms.
    It is easy to check that $F$ is full and faithful.
    Furthermore $F$ is surjective on the objects, because $A'$ and $A''$ are isomorph in $\cat{B}$.
    Thus $\cat{A} \eqv \cat{B}$.
\end{answer}

\begin{exercise}
        Let $\cat{A}$ and $\cat{B}$ be categories.  Prove that
        $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
\end{exercise}

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like 'God-given' or
  'defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by 'choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)] $\lambda\, a\, \lambda\,f.\, f\,a$
        \item[(b)] $\lambda\, (a, f).\, f\,a$
    \end{itemize}
\end{answer}

\begin{definition}{Product of Morphisms inside a Category}
  Let $\cat{A}$ be a category and $A, B, C, D \in \ob(\cat{A})$ with $f \from A \to C$ and $g \from B \to D$.
  The product morphism of $f$ and $g$, written $f \times g$, is a function
  \begin{align*}
    A \times B &\qquad\to\qquad C \times D\\
    (a, b) &\qquad\mapsto\qquad (f(a), g(b))
  \end{align*}
\end{definition}

\newcommand{\eval}[1]{\ensuremath{\mathsf{eval}^{#1}\xspace}}

\begin{exercise}
  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
  a functor $F_A:\mathbf{Set} \to \mathbf{Set}$ as follows:
  \begin{align*}
    F_A(B) & =  B^A \times A\\
    F_A(f: B \to C) & =  (f \circ \_) \times 1_A
  \end{align*}
  The function $\eval{A} \from F_A \to 1_{\mathbf{Set}}$, defined as $\eval{A}(f,a) := f(a)$, is a natural transformation if the following diagram commutes for each $g \from B \to C$.
\[
  \begin{tikzcd}
    F_A(B) = B^A \times A \arrow{rrr}{F_A(g)=(g\circ \_) \times 1_A} \arrow{dd}[swap]{\eval{A}_B} & & & 1_{\mathbf{Set}}(B) = B \arrow{dd}{\eval{A}_C} \\
                                                                                           & & & \\
    F_A(C) = C^A \times A \arrow{rrr}[swap]{1_{\mathbf{Set}}(g) = g} & & & 1_{\mathbf{Set}}(C) = C
  \end{tikzcd}
\]

Check that this diagram commutes.
\end{exercise}

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
            \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
            \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            By definition of natural isomorphism we have a natural transformation $\beta: G \Rightarrow F$, such that
            \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
            We show that $\forall A \in \cat{A}.\, \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}$.
            Suppose $A \in \cat{A}$.
            We know by definition of the identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
            The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of composition of natural transformations.
            $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
        \item[(ab)]
            We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
            \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}.\]
            First we prove that $\beta$ is a natural transformation.
            Suppose $f\from A \to A'$.
            We know that $\alpha$ is a natural transformation:
            \begin{align*}
                & G(f) \circ \alpha_A && = \alpha_{A'} \circ F(f)\\
                \Rightarrow\, & G(f) \circ \alpha_A \circ \beta_A && = \alpha_{A'} \circ F(f) \circ \beta_A\\
                \Rightarrow\, & \beta_{A'} \circ G(f) && = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
                \Rightarrow\, & \beta_{A'} \circ G(f) && = F(f) \circ \beta_A
            \end{align*}
            By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
            Therefore $\alpha$ is a natural isomorpism.
    \end{itemize}
\end{answer}

\begin{exercise}
A \emph{permutation} of a set $X$ is a bijection $X \to X$.
Write $\Sym(X)$ for the set of permutations of $X$.
A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
either $x \leq y$ or $y \leq x$.
Write $\Ord(X)$ for the set of total orders on $X$.

Let $\cat{B}$ denote the category of finite sets and bijections.

        \begin{itemize}
        \item[(a)]
Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
definitions should be canonical (no arbitrary choices).

        \item[(b)]
Show that there is no natural transformation $\Sym \to \Ord$.
(Hint: consider identity permutations.)

        \item[(c)]
For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
and $\Ord(X)$ have?
        \end{itemize}

Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
set $X$, there are exactly as many permutations of $X$ as there are total
orders on $X$, but there is no natural way of matching them up.)
\end{exercise}

\begin{exercise}
Let $F\from \cat{A} \to \cat{B}$ be a functor.

        \begin{itemize}
        \item[(a)]
Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
essentially surjective on objects.  (Hint: prove faithfulness before
fullness.)

        \item[(b)]
Now suppose instead that $F$ is full, faithful and essentially surjective
on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
and an isomorphism $\varepsilon\from F(G(B)) \to B$.  Prove that $G$ extends
to a functor in such a way that $(\varepsilon)_{B \in \cat{B}}$ is a natural
isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
        \end{itemize}
\end{exercise}

\begin{exercise}
This exercise makes precise the idea that linear algebra can equivalently
be done with matrices or with linear maps.

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F: \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && \forall n \in \NN\\
        F(A) & := \lambda\, x. \, A \cdot x && \forall n \times m \text{ matrices } A
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B$ be two $n \times m$ matrices over $K$ such that $F(A) = F(B)$.
            By using the standard basis in $K^m$ we get, that each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f: K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is surjective on the objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: