### 2017-05-26 sheets/talk-4/sheet4.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}
\[\uebunghead{Talk 4}{Galois Connections and Adjunctions}
\author {Sarah Mameche}
\section{Galois connections}

%Solutions preliminary

\begin{exercise}
Let $\mathscr{P} = (P, \preceq_P)$ and $\mathscr{Q} = (Q, \preceq_Q)$ be partially ordered sets, and $F \from P \to Q, G \from Q \to P$ monotone functions with $F \dashv G$.
\begin{enumerate}
\item Show that the following characterizations of the Galois connection are equivalent.
\begin{enumerate}
\item $p \preceq_P G (F (p)) \land q \preceq_Q F(G (q))$
\item $p \preceq_P G(q) \Leftrightarrow q \preceq_Q F(p)$
\end{enumerate}
\item Show that the following property holds for $F \dashv G$. 
\[  
\begin{aligned}
 \forall q \in Q.\; G(q) = min\{p \in P \such F(p) \preceq_Q q\}
\end{aligned}
\]
Formulate a similar statement $ \forall p \in P, F(p)$.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item ...
\item %For some $q \in Q$, set $S = \{p \in P \such F(p) \preceq_Q q}.
Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
\end{enumerate}
\end{answer}

\begin{exercise}
Consider the posets $\mathscr{N} = (\mathbb{N}, \leqslant), \mathscr{Q} = (\mathbb{Q}^+, \leqslant)$ 
and the inclusion function $I : \mathscr{N} \to  \mathscr{Q}$.
\begin{enumerate}
\item Find Galois connections $H \dashv I$ and $I \dashv G$.
\item Show (with the example from this exercise or a new one) that in general, Galois connections are not symmetric.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate} \item right adjunct: floor function  $\mathcal{F} : \mathcal{Q} \to  \mathcal{N}$ \\
left adjunct: ceiling function $\mathcal{C} : \mathcal{Q} \to  \mathcal{N}$
\item There can be no galois connection $F \dashv G$ from $\mathscr{Q}$ to $\mathscr{N} . F(q)=1 \Leftrightarrow 1 q<2,$ so $\{q\in \mathbb{Q}^+ \such F(q) = 1\} $ has no maximum. The statement follows with Ex.1b
\end{enumerate}\end{answer}

\section{Adjunctions}
\begin{exercise}
Recall the naturality axiom that has to hold for the isomorphism $\mathscr{B}(F(A), B) \; \cong \; \mathscr{A}(A, G(B)) $:

\begin{enumerate}[label=(\roman*)]
\item for any $g \from F(A) \to B$ and $q \from B \to B': \overline{q \of g} = G(q) \of \bar{g} $
\item for any $f \from A \to G(B) $ and $ p \from A' \to A: \overline{f \of p} = \bar{f} \of F(p)$
\end{enumerate}
Show that (i) and (ii) can be replaced by (iii) $\forall f, q, p. \;\overline{G(q) \of f \of p} = q \of \bar{f} \of F(p)$
\end{exercise}

\begin{answer}
(i), (ii) $\Rightarrow$ (iii)\\
composing any $q \from B \to B'$ with (ii): $\forall f, p: \forall q: q \of \bar{f} \of F(p) = q \of \overline{f \of p} $ (by ii). \\Since $f \from p \to F(A') \to B$, its transpose $h=:\overline{f \of p} \from A' \to B.$ \\So from (i) and quantif. over A: $q \of \overline{f \of p}  = q \of h = \overline{G(q) \of \bar{h}} = \overline{G(q) \of f \of p}$\\
(iii) $\Rightarrow$ (i)\\
%(iii) quantifies over all $p$, so set $p = id_A$ (exists for each object). Then $f \of p \from A \to G(B)$, and $\bar{f} \of F(p) \from F(A) \to B.$ 
Assume (i) is false. 
Taking a $g \from F(A) \to B$, you get $h = g  \of id_B  \from F(A) \to B$. 
Then the equality (i) does not hold for h, that is \\ 
$\forall q, \overline{q \of h} \not= G(q) \of \bar{h}$ 
\\$ \Leftrightarrow \overline{q \of g  \of id_B} \not= G(q)  \of \overline{h}$ ( Def h)\\
$\Leftrightarrow q \of g  \of id_B \not= \overline{G(q)  \of \overline{h}}$ (Def transposition)

$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{h}} $  (define f as $\bar {g}.$ Also, $  F(id_A) = id_B$ by Def of functors)
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{g}} \; (h = g \of id_B = g)$
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of f \of id_A}\;  (f = \bar{g} $ by def. and $f = f \of id_A)$
\\So (iii) is false for $f = \bar {g}$ and $ p= id_A$. \\(iii)  $\Rightarrow$ (i) similarly.
\end{answer}

\begin{exercise}
Find a right adjunct to the forgetful functor $F \from $\textbf{Set }$ \to $\textbf{ Rel}.
\end{exercise}
\begin{answer}
\end{answer}

\begin{exercise}
\begin{enumerate}
\item Find a functor that is both left and right adjoint to another functor.
\item Find a functor that is self-adjoint.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item We have seen that initial (terminal) objects of categories correspond to left (right) adjoints. So for any category that has a zero object, there is a functor that is both left and right adjoint. Also, an equivalence between two categories is both a left and right adjoint.
\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex. 7) is self-adjoint. Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus to functions $B \to \mathcal{P}(A)$.
\end{enumerate}
\end{answer}

\begin{exercise}
Show that adjunctions preserve initial objects. I.e., if  
\begin{tikzcd}[ampersand replacement=\&, every label/.append style = {font = \footnotesize}]
\mathscr{A} \arrow[r, shift left,"F"] 
\& \mathscr{B} \arrow[l, shift left, "G"] 
\end{tikzcd}
with $F\dashv G$ and $I$ is initial in $ \mathscr{A}$, then $F(I)$ is initial in $\mathscr{B}$. Dualize the statement.
\end{exercise}

\begin{answer}
Assume $I$ is initial in $ \mathscr{A}$, then maps $I \to A$ are unique $\forall A \in A$. This implies that maps  $I \to G(B)$ are unique $\forall B$ because $G(B) \in ob(\mathscr{A})$. From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique $\forall B$, so $F(I)$ initial in $\mathscr{B}$.\\ It can be shown in the same way that terminal objects are preserved by $G$.\end{answer}

\begin{exercise}
For sets $X, Y$ and a function $f: X \to Y$, show that the following operations are adjunctions:\\
$ \forall A \in \mathcal{P}(X).\; F \from \mathcal {P}(X) \to \mathcal {P}(Y ) $ with $ F(A) := \{f(a) \such a \in A \}$ and\\
$\forall B\in \mathcal{P}(Y). \; F^{-1} \from \mathcal {P}(Y) \to \mathcal {P}(X)$ with $ F^{-1}(B):=\{a\in X \such f(a)\in B\}$
\\Adapt the construction so that it works for groups (alternatively topological spaces) instead of sets. 
\end{exercise}

\begin{answer}
$F \dashv F^{-1}$, meaning $F(f)(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}$.\\
Assume $F \dashv F^{-1}$. Then $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}$\\
Assume $X \subseteq F^{-1}$ and take $b \in F^{-1}$. By Def $f(x) = b$ for some x, and with $X \subseteq F^{-1}, b = f(x) \in Y$\\
Groups: f homomorphism, $\mathcal{P}(G)$ set of subgroups, same proof \\
Top. Sp: continuous maps and subspaces

\end{answer}

\begin{exercise}
Consider a functor $Ob \from $\textbf {CAT }$  \to $\textbf { Set} which maps a category to its set of objects, and a functor to the mapping induced on objects. Are there left and right adjoints?
\end{exercise}

\begin{answer}
There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$. There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$.
\end{answer}

\begin{exercise}
Let $G$ be a group. \begin{enumerate}
\item Find functors between \textbf{Set} and the functor category $[G,$ \textbf{Set}$]$ of left $G$-sets. Which adjunnctions are there between them?
\item Similarly, find functors \textbf{Vect}$_k$ and $[G,$ \textbf{Vect}$_k]$ and look for adjunctions.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item Objects of $[G,$ \textbf{Set}$]$ are functors $F_S$, with one object (set S) and with maps $f_g \from S\to S$ (for each $g \in G$, left group action). Maps in  $[G,$ \textbf{Set}$]$ are natural transformations between the functors. The canonical functor $F \from$ \textbf{Set}$ \to [G,$ \textbf{Set}$]$ maps a set $S$ to the functor with $S$ as the underlying set. Functions $S\to S' $ are mapped to the natural transformation between the resp. functors $F_S, F_S'$. A functor G which goes back to \textbf{Set} is defined similarly. Then $F \dashv G$ and $G \dashv F$, because every function $S \to S'$ corresponds by construction to exactly one natural transformation $\alpha \from F_S \to F_S'$. The correspondence is natural in S, S' because $\alpha$ preserves the group action: $\alpha (g . s) = g . (\alpha . s)\; \forall s \in S$ (because $\alpha$ is a natural transformation).

\item The Functors $F_V$ in  $[G,$ \textbf{Vect}$_k]$ select a vector space $V$, and have linear maps $f_g: V \to W$ for each $g \in G$. Construct $F$ and $G$ as in (a), then  $F \dashv G$ and $G \dashv F$, because the natural transformations preserve the group action.
\end{enumerate}
\end{answer}

\]
\end{document}