### 2017-05-26 sheets/basis-3/sheet-basis-3.tex
Partialy Solution for Basis 3.7

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}

\begin{hint}
  Read Chapter 1.3.
\end{hint}

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}
\begin{answer}
    \begin{minipage}{0.5\textwidth}
        $\cat{A}$:
        \[
            \begin{tikzcd}
                A \arrow{d}{f}\\
                B
            \end{tikzcd}
        \]
    \end{minipage}
    \begin{minipage}{0.5\textwidth}
        $\cat{B}$:
        \[
            \begin{tikzcd}
                A' \arrow{dr}{f'} \arrow[bend left]{rr}{g} &    & A'' \arrow{dl}{f''} \arrow{ll}{g^{-1}} \\
                                                           & B' &
            \end{tikzcd}
        \]
    \end{minipage}
    The identity morphisms have been left out for simplicity.
    First of all note that there is no bijection between the objects of $\cat{A}$ and $\cat{B}$, so $\cat{A} \not\iso \cat{B}$.\\
    We now show $\cat{A} \eqv \cat{B}$.
    Therefore we define a functor $F$:
    \begin{align*}
        F(A) := A'\\
        F(B) := B'\\
        F(f) := f'
    \end{align*}
    In addition $F$ maps the identity morhisms to their respective identity morphisms.
    It is easy to check that $F$ is full and faithful.
    Furthermore $F$ is essentially surjective on objects because $A'$ and $A''$ are isomorphic in $\cat{B}$.
    Thus $\cat{A} \eqv \cat{B}$.
\end{answer}

\begin{exercise}
        Let $\cat{A}$ and $\cat{B}$ be categories. Prove that
        $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
\end{exercise}

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like `God-given' or
  `defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by `choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)] $\lambda\, a.\, \lambda\,f.\, f\,a$
        \item[(b)] $\lambda\, (a, f).\, f\,a$
    \end{itemize}
\end{answer}

\begin{definition}{Product of Morphisms inside a Category}
  Let $\cat{A}$ be a category and $A, B, C, D \in \ob(\cat{A})$ with $f \from A \to C$ and $g \from B \to D$.
  The product morphism of $f$ and $g$, written $f \times g$, is a function
  \begin{align*}
    A \times B &\qquad\to\qquad C \times D\\
    (a, b) &\qquad\mapsto\qquad (f(a), g(b))
  \end{align*}
\end{definition}

\newcommand{\eval}[1]{\ensuremath{\mathsf{eval}^{#1}\xspace}}

\begin{exercise}
  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
  a functor $F_A\from \mathbf{Set} \to \mathbf{Set}$ as follows:
  \begin{align*}
    F_A(B) & =  B^A \times A\\
    F_A(f: B \to C) & =  (f \circ \_) \times 1_A
  \end{align*}
  The function $\eval{A} \from F_A \to 1_{\mathbf{Set}}$, defined as $\eval{A}(f,a) := f(a)$, is a natural transformation if the following diagram commutes for each $g \from B \to C$.
\[
  \begin{tikzcd}
    F_A(B) = B^A \times A \arrow{rrr}{F_A(g)=(g\circ \_) \times 1_A} \arrow{dd}[swap]{\eval{A}_B} & & & 1_{\mathbf{Set}}(B) = B \arrow{dd}{\eval{A}_C} \\
                                                                                           & & & \\
    F_A(C) = C^A \times A \arrow{rrr}[swap]{1_{\mathbf{Set}}(g) = g} & & & 1_{\mathbf{Set}}(C) = C
  \end{tikzcd}
\]

Check that this diagram commutes.
\end{exercise}
\begin{answer}
  Let $(x,y) \in B^A \times A$. We have to show that $\eval{A}_C (F_A(g) (x,y)) = 1_{\mathbf {Set}}(g) (\eval{A}_B (x, y))$.
  By unfolding the definitions of $\times$, $\eval{A}$ and $F_A$ we get:
  \[
    \eval{A}_C (F_A(g) (x,y)) =
    \eval{A}_C ((g \circ \_)\,x, 1_A\,y) =
    \eval{A}_C ((g \circ x, y)) =
    g (x (y))
    \]
  By unfolding the definitions of $1_{\mathbf {Set}}$ and $\eval{A}$ we get:
  \[
    1_{\mathbf {Set}}(g) (\eval{A}_B (x, y)) =
    g (\eval{A}_B (x, y)) =
    g (x (y))
    \]
\end{answer}

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
            \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
            \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            By definition of natural isomorphism we have a natural transformation $\beta\from G \to F$, such that
            \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
            We show that $\forall A \in \cat{A}.\, \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}$.
            Suppose $A \in \cat{A}$.
            We know by definition of the identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
            The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of composition of natural transformations.
            $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
        \item[(b)]
            We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
            \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}.\]
            First we prove that $\beta$ is a natural transformation.
            Suppose $f\from A \to A'$.
            We know that $\alpha$ is a natural transformation:
            \begin{alignat*}{2}
                & & G(f) \circ \alpha_A & = \alpha_{A'} \circ F(f) \\
                &\Rightarrow &\quad G(f) \circ \alpha_A \circ \beta_A & = \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & G(f) & = \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & \beta_{A'} \circ G(f) & = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & \beta_{A'} \circ G(f) & = F(f) \circ \beta_A
            \end{alignat*}
            By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
            Therefore $\alpha$ is a natural isomorpism.
    \end{itemize}
\end{answer}

\begin{exercise}
A \emph{permutation} of a set $X$ is a bijection $X \to X$.
Write $\Sym(X)$ for the set of permutations of $X$.
A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
either $x \leq y$ or $y \leq x$.
Write $\Ord(X)$ for the set of total orders on $X$.

Let $\cat{B}$ denote the category of finite sets and bijections.

        \begin{itemize}
        \item[(a)]
Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
definitions should be canonical (no arbitrary choices).

        \item[(b)]
Show that there is no natural transformation $\Sym \to \Ord$.
(Hint: consider identity permutations.)

        \item[(c)]
For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
and $\Ord(X)$ have?
        \end{itemize}

Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
set $X$, there are exactly as many permutations of $X$ as there are total
orders on $X$, but there is no natural way of matching them up.)
\end{exercise}

\begin{answer}
  \begin{itemize}
    \item[(a)]
      $\Sym(f) = \lambda g.f \circ g$, where $g$ is a permutation\\
      $\Ord(f) = \lambda (a, b).(f(a), f(b))$, where $a, b \in X$
    \item[(b)]
      Let's consider a non-identity permutation $h$. We can build an ordering
      from it by taking $(x, h(x))$ as an element of ordering relation for
      $\forall x \in \text{ some finite set } X$, then filling the gaps by
      reflexivity and transitivity.

      However, it's not the case for an identity permutation $i$, as for that
      permutation the $i(x) = x$ which leaves us only with reflexivity pairs if
      we try to build an ordering from that bijection. Thus, we cannot define
      formation of an ordering from an identity permutation and therefore
      there's no natural transformation from $\Sym$ to $\Ord$.
    \item[(c)]
      $|\Sym(X)| = n$!\\
      $|\Ord(X)| = n$!
  \end{itemize}
  We have seen that there's no natural transformation $\Sym \to \Ord$.
  Nonetheless, we have also seen that they have the same number of
  elements and an isomorphism between them can actually be formed: it's not
  hard to construct all permutations or orderings of some finite set knowing
  its elements.

  This leads us to the conclusion that $\Sym(X) \iso \Ord(X)$ for all $X \in
  \cat{B}$, but not \emph{naturally} in $X \in \cat{B}$.
\end{answer}

\begin{exercise}
Let $F\from \cat{A} \to \cat{B}$ be a functor.

        \begin{itemize}
        \item[(a)]
Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
essentially surjective on objects.  (Hint: prove faithfulness before
fullness.)

        \item[(b)]
Now suppose instead that $F$ is full, faithful and essentially surjective
on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
and an isomorphism $\epsln_B\from F(G(B)) \to B$.  Prove that $G$ extends
to a functor in such a way that $(\epsln)_{B \in \cat{B}}$ is a natural
isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            Since $F$ is an equivalance there is a functor $G\from \cat{B} \to \cat{A}$ and a natural isomorphism $\eta\from 1_{\cat{A}} \to G \circ F$.
            So the following diagram commutes for all $f\from A \to A'$:
            \[
                \begin{tikzcd}
                    A \arrow{r}{f} \arrow{d}{\eta_A} & A' \arrow{d}{\eta_{A'}} \\
                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
                \end{tikzcd}
            \]
            Also we have a natural isomorphism $\epsln\from F \circ G \to 1_{\cat{B}}$ such that for all $g\from B \to B'$ the diagram commutes:
            \[
                \begin{tikzcd}
                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
                    B \arrow{r}{g}  & B'
                \end{tikzcd}
            \]
            \begin{description}
                \item[$F$ is faithful:]
                    Let $A, A' \in \cat{A}$ and $f,f'\from A \to A'$ such that $F(f) = F(f')$.
                    Using this commutativity for $\eta$ we obtain our goal.
                    \begin{align*}
                        f & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f\\
                          & = \eta_{A'}^{-1} \circ G(F(f)) \circ \eta_A\\
                          & = \eta_{A'}^{-1} \circ G(F(f')) \circ \eta_A\\
                          & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f'\\
                          & = f'
                    \end{align*}
                \item[$F$ is full:]
                    %TODO
                \item[$F$ is essentially surjective on objects:]
                    Let $B \in \cat{B}$.
                    Then $F(G(B)) \iso B$ by $\epsln_B$.
            \end{description}
        \item[(b)]
            $F$ is essentially surj. on obj. so $\forall B \in \cat{B}.\, \exists A \in \cat{A}.\, F(A) \iso B$.
            We denote by $h(A, B) \from F(A) \to B$ this isomorphism.
            For each $B \in \cat{B}$ we define $G(B)$ such that $F(G(B)) \iso B$ under $h(G(B), B)$ and $\epsln_B\from F(G(B)) \to B := h(G(B), B)$.
            Let $g\from B \to B'$ be a morphism in $\cat{B}$.
            Then $g' := \epsln_{B'}^{-1} \circ g \circ \epsln_{B} \from F(G(B)) \to F(G(B'))$ is a morphism in the image of $F$.
            We extend $G$ to a functor by mapping each morphism $g\from B \to B'$ to the morphism $f\from G(B) \to G(B')$ such that $g' = F(f)$.
            This morphism $f$ exists and is unique because $F$ is full and faithful.
            We have to show that $\epsln$ is a natural transformation.
            \[
                \begin{tikzcd}
                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
                    B \arrow{r}{g}                                   & B'
                \end{tikzcd}
            \]
            The diagram commutes since $F(G(g)) = g' = \epsln_{B'}^{-1} \circ g \circ \epsln_{B}$.\\
            Next we define $\eta_A$ such that $F(\eta_A) = \epsln_{F(A)}^{-1}$.
            Again this is possible because $\epsln_{F(A)}^{-1}\from F(A) \to F(G(F(A)))$ is in the image of $F$.
            \[
                \begin{tikzcd}
                    A \arrow{d}{\eta_A} \arrow{r}{f} & A' \arrow{d}{\eta_{A'}} \\
                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
                \end{tikzcd}
            \]
            \begin{align*}
                F(G(F(f))) & = \epsln_{F(A')}^{-1} \circ F(f) \circ \epsln_{F(A)}\\
                           & = F(\eta_{A'}) \circ F(f) \circ F(\eta_A^{-1})\\
                           & = F(\eta_{A'} \circ f \circ \eta_A^{-1})
            \end{align*}
            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithfull.
            This shows that $\eta$ is a natural isomorphism.
            Therefore $F$ is an equivalence.
    \end{itemize}
\end{answer}

\begin{exercise}
This exercise makes precise the idea that linear algebra can equivalently
be done with matrices or with linear maps.

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F\from \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && n \in \NN\\
        F(A\from m \to n) & := \lambda\, x. \, A \cdot x && A \text{ is a } n \times m \text{ matrix}
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B\from m \to n$, that is $n \times m$ matrices over $K$, such that $F(A) = F(B)$.
            By using the standard basis $(e_j)_{j = 1}^m$ in $K^m$ we get:
            \[A \cdot e_j = F(A)(e_j) = F(B)(e_j) = B \cdot e_j \quad\forall 1 \le j \le m\]
            So each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f\from K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is essentially surjective on objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
    Thus $F$ is an equivalence and so $\Mt \eqv \FDVect_K$.
    The functor $F$ is even canonical since we didn't make arbitrary choices.
    But the functor from $\FDVect_K$ to $\Mt$ involves choosing a basis like in the case for showing that $F$ is full.
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: