### 2017-05-29 sheets/basis-3/sheet-basis-3.tex
WIP: Solution for 3.2

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\begin{document}

\author{Joachim Bard, Maximilian Wuttke, Nikita Ziuzin}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Basis 3}{Maps between Functors -- Natural Transformations}

\begin{hint}
  Read Chapter 1.3.
\end{hint}

\begin{exercise}
  Define categories $\cat{A}$ and $\cat{B}$ such that either $\cat{A} \iso \cat{B}$ in $\CAT$ or $\cat{A} \eqv \cat{B}$, but not both.
  Do not use the example given during the lecture.
\end{exercise}
\begin{answer}
    \begin{minipage}{0.5\textwidth}
        $\cat{A}$:
        \[
            \begin{tikzcd}
                A \arrow{d}{f}\\
                B
            \end{tikzcd}
        \]
    \end{minipage}
    \begin{minipage}{0.5\textwidth}
        $\cat{B}$:
        \[
            \begin{tikzcd}
                A' \arrow{dr}{f'} \arrow[bend left]{rr}{g} &    & A'' \arrow{dl}{f''} \arrow{ll}{g^{-1}} \\
                                                           & B' &
            \end{tikzcd}
        \]
    \end{minipage}
    The identity morphisms have been left out for simplicity.
    First of all note that there is no bijection between the objects of $\cat{A}$ and $\cat{B}$, so $\cat{A} \not\iso \cat{B}$.\\
    We now show $\cat{A} \eqv \cat{B}$.
    Therefore we define a functor $F$:
    \begin{align*}
        F(A) := A'\\
        F(B) := B'\\
        F(f) := f'
    \end{align*}
    In addition $F$ maps the identity morhisms to their respective identity morphisms.
    It is easy to check that $F$ is full and faithful.
    Furthermore $F$ is essentially surjective on objects because $A'$ and $A''$ are isomorphic in $\cat{B}$.
    Thus $\cat{A} \eqv \cat{B}$.
\end{answer}

\begin{exercise}
        Let $\cat{A}$ and $\cat{B}$ be categories. Prove that
        $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$.
\end{exercise}

\begin{answer}
  We have to give two functors
  $H \from \ftrcat{\cat{A}^\op}{\cat{B}^\op} \to \ftrcat{\cat{A}}{\cat{B}}^\op$
  and
  $J \from \ftrcat{\cat{A}}{\cat{B}}^\op \to \ftrcat{\cat{A}^\op}{\cat{B}^\op}$
  such that $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.

  First we define $H$. For that we first map objects of
  $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$ (being itself functors from $\cat{A}^\op$ to $\cat{B}^\op$)
  to objects of $\ftrcat{\cat{A}}{\cat{B}}^\op$ (being itself functors from $\cat{A}$ to $\cat{B}$).

  Let $F\from \cat{A}^\op \to \cat{B}^\op$ be an object of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
  Because every object $A : \cat{A}^\op$ is also an object of $\cat{A}$ we can define
  $H(F)(A) := F(A)$.
  Now let $f \from A \to B$ be an morphism in $\cat{A}^\op$. We have $F(f^\op)=g^\op$ for some morphism $g$ in $\cat{B}^\op$.
  We define $H(F)(f) := F(f^\op)^\op = g$.

  Now let $F, G \from \cat{A}^\op \to \cat{B}^\op$ be objects and
  $\alpha \from F \to G$ be an morphism of $\ftrcat{\cat{A}^\op}{\cat{B}^\op}$.
  $\alpha = (F(A) \stackrel{\alpha_A}{\to} G(A))_A$ is an natural transformation.
  We define $H(\alpha) := \beta \from H(G) \to H(F) := (\alpha_A^\op)_A$ and have to show that $\beta$ is indeed such a natural transformation.
  Let $A, A' : \cat{A}$ and $f \from A \to A'$. We have to show that the following diagram commutes:
  \[
    \begin{tikzcd}
      H(G)(A) \arrow{r}{H(G)(f)} \arrow{d}{H(\alpha)_A} & H(G)(A') \arrow{d}{H(\alpha)_A'} \\
      H(F)(A) \arrow{r}{H(F)(f)} & H(F)(A')
    \end{tikzcd}
  \]
  By unfolding the definition of $H$ this diagram commutes if and only if the following diagram commutes:
  \[
    \begin{tikzcd}
      G(A) & G(A') \arrow{l}{G(f^\op)} \\
      F(A) \arrow{u}{\alpha_A} & F(A') \arrow{l}{F(f^\op)} \arrow{u}{\alpha_{A'}}
    \end{tikzcd}
  \]
  This diagram commutes because $\alpha$ is a natural transformation.

  Secondly we define $J$ using a similar construction as above and get
  $H \circ J = 1_{\ftrcat{\cat{A}}{\cat{B}}^\op}$ and $J \circ H = 1_{\ftrcat{\cat{A}^\op}{\cat{B}^\op}}$.

  Now by definition of $\iso$,
  $\ftrcat{\cat{A}^\op}{\cat{B}^\op} \iso \ftrcat{\cat{A}}{\cat{B}}^\op$
  holds.
\end{answer}

\begin{definition}{Canonical according to Leinster}
  "Canonical is an informal word, meaning something like `God-given' or
  `defined without making arbitrary choices'.  For example, for any two sets
  $A$ and $B$, there is a canonical bijection $A \times B \to B \times A$
  defined by $(a, b) \mapsto (b, a)$, and there is a canonical function $A
  \times B \to A$ defined by $(a, b) \mapsto a$.  But the function $B \to A$
  defined by `choose an element $a_0 \in A$ and send everything to $a_0$' is
  not canonical, because the choice of $a_0$ is arbitrary."
\end{definition}

\begin{exercise}
  Let $A$ and $B$ be sets, and denote by $B^A$ the set of functions from $A$
  to $B$.  Write down:
  \begin{itemize}
  \item[(a)] a canonical function $A \to B^{(B^A)}$,
  \item[(b)] a canonical function $A \times B^{A} \to B$.
  \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)] $\lambda\, a.\, \lambda\,f.\, f\,a$
        \item[(b)] $\lambda\, (a, f).\, f\,a$
    \end{itemize}
\end{answer}

\begin{definition}{Product of Morphisms inside a Category}
  Let $\cat{A}$ be a category and $A, B, C, D \in \ob(\cat{A})$ with $f \from A \to C$ and $g \from B \to D$.
  The product morphism of $f$ and $g$, written $f \times g$, is a function
  \begin{align*}
    A \times B &\qquad\to\qquad C \times D\\
    (a, b) &\qquad\mapsto\qquad (f(a), g(b))
  \end{align*}
\end{definition}

\newcommand{\eval}[1]{\ensuremath{\mathsf{eval}^{#1}\xspace}}

\begin{exercise}
  For a fixed set $A$, the map taking $B$ to $B^A \times A$ can be extended to
  a functor $F_A\from \mathbf{Set} \to \mathbf{Set}$ as follows:
  \begin{align*}
    F_A(B) & =  B^A \times A\\
    F_A(f: B \to C) & =  (f \circ \_) \times 1_A
  \end{align*}
  The function $\eval{A} \from F_A \to 1_{\mathbf{Set}}$, defined as $\eval{A}(f,a) := f(a)$, is a natural transformation if the following diagram commutes for each $g \from B \to C$.
\[
  \begin{tikzcd}
    F_A(B) = B^A \times A \arrow{rrr}{F_A(g)=(g\circ \_) \times 1_A} \arrow{dd}[swap]{\eval{A}_B} & & & 1_{\mathbf{Set}}(B) = B \arrow{dd}{\eval{A}_C} \\
                                                                                           & & & \\
    F_A(C) = C^A \times A \arrow{rrr}[swap]{1_{\mathbf{Set}}(g) = g} & & & 1_{\mathbf{Set}}(C) = C
  \end{tikzcd}
\]

Check that this diagram commutes.
\end{exercise}
\begin{answer}
  Let $(x,y) \in B^A \times A$. We have to show that $\eval{A}_C (F_A(g) (x,y)) = 1_{\mathbf {Set}}(g) (\eval{A}_B (x, y))$.
  By unfolding the definitions of $\times$, $\eval{A}$ and $F_A$ we get:
  \[
    \eval{A}_C (F_A(g) (x,y)) =
    \eval{A}_C ((g \circ \_)\,x, 1_A\,y) =
    \eval{A}_C ((g \circ x, y)) =
    g (x (y))
    \]
  By unfolding the definitions of $1_{\mathbf {Set}}$ and $\eval{A}$ we get:
  \[
    1_{\mathbf {Set}}(g) (\eval{A}_B (x, y)) =
    g (\eval{A}_B (x, y)) =
    g (x (y))
    \]
\end{answer}

\begin{exercise}
        Let $\xymatrix@1{\cat{A}\rtwocell^F_G{\alpha} &\cat{B}}$ be a natural transformation.
        \begin{itemize}
            \item[(a)] Suppose $\alpha$ is a natural isomorphism.
                Show that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all $A \in \cat{A}$.
            \item[(b)] Now assume that $\alpha_A\from F(A) \to G(A)$ is an isomorphism for all
                $A \in \cat{A}$. Prove that $\alpha$ is a natural isomorphism.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            By definition of natural isomorphism we have a natural transformation $\beta\from G \to F$, such that
            \[\beta \circ \alpha = 1_{F} \text{ and } \alpha \circ \beta = 1_{G}. \]
            We show that $\forall A \in \cat{A}.\, \beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}$.
            Suppose $A \in \cat{A}$.
            We know by definition of the identity natural transformation that $(\beta \circ \alpha)_A = 1_{F(A)}$.
            The claim $\beta_A \circ \alpha_A = 1_{F(A)}$ follows by definition of composition of natural transformations.
            $\alpha_A \circ \beta_A = 1_{G(A)}$ is shown analogously.
        \item[(b)]
            We know that for all $A \in \cat{A}$ there is a $\beta_A$ such that
            \[\beta_A \circ \alpha_A = 1_{F(A)} \text{ and } \alpha_A \circ \beta_A = 1_{G(A)}.\]
            First we prove that $\beta$ is a natural transformation.
            Suppose $f\from A \to A'$.
            We know that $\alpha$ is a natural transformation:
            \begin{alignat*}{2}
                & & G(f) \circ \alpha_A & = \alpha_{A'} \circ F(f) \\
                &\Rightarrow &\quad G(f) \circ \alpha_A \circ \beta_A & = \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & G(f) & = \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & \beta_{A'} \circ G(f) & = \beta_{A'} \circ \alpha_{A'} \circ F(f) \circ \beta_A\\
                &\Rightarrow & \beta_{A'} \circ G(f) & = F(f) \circ \beta_A
            \end{alignat*}
            By similar reasoning to (a) we get that $\beta$ is the inverse of $\alpha$.
            Therefore $\alpha$ is a natural isomorpism.
    \end{itemize}
\end{answer}

\begin{exercise}
A \emph{permutation} of a set $X$ is a bijection $X \to X$.
Write $\Sym(X)$ for the set of permutations of $X$.
A \emph{total order} on a set $X$ is an order $\leq$ such that for all $x, y \in X$,
either $x \leq y$ or $y \leq x$.
Write $\Ord(X)$ for the set of total orders on $X$.

Let $\cat{B}$ denote the category of finite sets and bijections.

        \begin{itemize}
        \item[(a)]
Give a definition of $\Sym$ on maps in $\cat{B}$ in such a way that $\Sym$
becomes a functor $\cat{B} \to \Set$.  Do the same for $\Ord$.  Both your
definitions should be canonical (no arbitrary choices).

        \item[(b)]
Show that there is no natural transformation $\Sym \to \Ord$.
(Hint: consider identity permutations.)

        \item[(c)]
For an $n$-element set $X$, how many elements do the sets $\Sym(X)$
and $\Ord(X)$ have?
        \end{itemize}

Conclude that $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$, but not
\emph{naturally} in $X \in \cat{B}$.  (The moral is that for each finite
set $X$, there are exactly as many permutations of $X$ as there are total
orders on $X$, but there is no natural way of matching them up.)
\end{exercise}

\begin{answer}
  In the following we will consider a representation of the permutation as a
  tuple of two elements $(a, b)$, where $a \mapsto b$ is an element of a
  bijection.
  \begin{itemize}
    \item[(a)]
      For any $A \in ob(\cat{B}), C \in ob(\Set), f: A \to C$
      \begin{align*}
        \Sym(f) &= \lambda (a, b). (f(a), f(b))\\
        \Ord(f) &= \lambda (a, b).(f(a), f(b))
      \end{align*}
    \item[(b)]
      Let's fix a finite set $A$ and define $\Sym \to \Ord$ on it.\\
      \begin{align*}
        A &= \{a, b\}\\
        \Sym(A) &= \{\{(a, a), (b, b)\}, \{(a, b), (b, a)\}\}\\
        \Ord(A) &= \{\{(a, b), (a, a), (b, b)\}, \{(b, a), (a, a), (b, b)\}\},\\
      \end{align*}
      Natural transformation $\alpha: \Sym(A) \to \Ord(A)$:\\
      \begin{align*}
        \alpha_A(\{(a, a), (b, b)\}) &= \{(a, b), (a, a), (b, b)\}\\
        \alpha_A(\{(a, b), (b, a)\}) &= \{(b, a), (a, a), (b, b)\}
      \end{align*}
      And a permutation of $A$:
      \begin{align*}
        f(a) &= b\\
        f(b) &= a
      \end{align*}
      Now, for $\alpha$ to indeed be a natural transformation, the  naturality
      axiom should hold for all arrows $A \to A$, including $f$. Let's check if
      it holds by applying $f$.\\
      \begin{align*}
        &\alpha_A(f(\{(a, a), (b, b)\})) &= \alpha_A(\{(b, b), (a, a)\}) =&& \{(b, a), (b, b), (a, a)\}\\
        &\alpha_A(f(\{(a, b), (b, a)\})) &= \alpha_A(\{(b, a), (a, b)\}) =&& \{(a, b), (b, b), (a, a)\}
      \end{align*}
      Here $\alpha_A$ contradicts our definition of it and we can conclude that
      it is not a natural transformation. The only other possibility of
      defining $\alpha_A$ would be to switch the images of sets in the
      definition, but the applying $f$ will yield us a similar contradictory
      result.  From this we see that no matter how we define $\alpha_A$, the
      naturality axiom will not hold for our $f$, hence we cannot have a
      natural transformation for $\Sym(A) \to \Ord(A)$ and, in general, a
      natural transformation $\Sym \to \Ord$.
    \item[(c)]
      From combinatorics we know that the number of all permutations of a
      finite set of size $n$ is exactly $n$!:\\
      \begin{equation*}
        |\Sym(X)| = n!
      \end{equation*}
      We can also notice that we may consider permutations as an order too,
      hence the number of all orders is:
      \begin{equation*}
        |\Ord(X)| = n!
      \end{equation*}
  \end{itemize}
  So we know that $\Sym$ and $\Ord$ have exactly the same number of elements
  and generally we can come up with a mapping between those elements, such that
  $\Sym(X) \iso \Ord(X)$ for all $X \in \cat{B}$. However, we have also shown
  that we cannot define a natural transformation $\Sym \to \Ord$ which gives
  that there's no natural isomorphism between those categories.
\end{answer}

\begin{exercise}
Let $F\from \cat{A} \to \cat{B}$ be a functor.

        \begin{itemize}
        \item[(a)]
Suppose that $F$ is an equivalence.  Prove that $F$ is full, faithful and
essentially surjective on objects.  (Hint: prove faithfulness before
fullness.)

        \item[(b)]
Now suppose instead that $F$ is full, faithful and essentially surjective
on objects.  For each $B \in \cat{B}$, choose an object $G(B)$ of $\cat{A}$
and an isomorphism $\epsln_B\from F(G(B)) \to B$.  Prove that $G$ extends
to a functor in such a way that $(\epsln_B)_{B \in \cat{B}}$ is a natural
isomorphism $F \circ G \to 1_{\cat{B}}$.  Then construct a natural isomorphism
$1_{\cat{A}} \to G \circ F$, thus proving that $F$ is an equivalence.
        \end{itemize}
\end{exercise}
\begin{answer}
    \begin{itemize}
        \item[(a)]
            Since $F$ is an equivalance there is a functor $G\from \cat{B} \to \cat{A}$ and a natural isomorphism $\eta\from 1_{\cat{A}} \to G \circ F$.
            So the following diagram commutes for all $f\from A \to A'$:
            \[
                \begin{tikzcd}
                    A \arrow{r}{f} \arrow{d}{\eta_A} & A' \arrow{d}{\eta_{A'}} \\
                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
                \end{tikzcd}
            \]
            Also we have a natural isomorphism $\epsln\from F \circ G \to 1_{\cat{B}}$ such that for all $g\from B \to B'$ the diagram commutes:
            \[
                \begin{tikzcd}
                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
                    B \arrow{r}{g}  & B'
                \end{tikzcd}
            \]
            \begin{description}
                \item[$F$ is faithful:]
                    Let $A, A' \in \cat{A}$ and $f,f'\from A \to A'$ such that $F(f) = F(f')$.
                    Using the naturality axiom for $\eta$ we obtain our goal.
                    \begin{align*}
                        f & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f\\
                          & = \eta_{A'}^{-1} \circ G(F(f)) \circ \eta_A\\
                          & = \eta_{A'}^{-1} \circ G(F(f')) \circ \eta_A\\
                          & = \eta_{A'}^{-1} \circ \eta_{A'} \circ f'\\
                          & = f'
                    \end{align*}
                    By analogous reasoning using $\epsln$ we obtain that $G$ is also faithful.
                \item[$F$ is full:]
                    Let $A, A' \in \cat{A}$ and $g \from F(A) \to F(A')$.
                    We define $f := \eta_{A'}^{-1} \circ G(g) \circ \eta_A$ and use the naturality axiom for $\eta$ in order to get:
                    \begin{align*}
                        G(F(f)) & = \eta_A' \circ f \circ \eta_A^{-1}\\
                                & = \eta_A' \circ \eta_{A'}^{-1} \circ G(g) \circ \eta_A \circ \eta_A^{-1}\\
                                & = G(g)
                    \end{align*}
                    Since $G$ is faithful we obtain $F(f) = g$.
                    Thus $F$ is full.
                \item[$F$ is essentially surjective on objects:]
                    Let $B \in \cat{B}$.
                    Then $F(G(B)) \iso B$ by $\epsln_B$.
            \end{description}
        \item[(b)]
            For each $B \in \cat{B}$ the class of objects $A \in \cat{A}$ which fulfill $F(A) \iso B$ is non empty, because $F$ is essentially surjective on objects.
            We define $G(B)$ as such an $A$ using the axiom of choice.
            In addition we have an isomorphism $\epsln_B\from F(G(B)) \to B$.
            Let $g\from B \to B'$ be a morphism in $\cat{B}$.
            Then $g' := \epsln_{B'}^{-1} \circ g \circ \epsln_{B} \from F(G(B)) \to F(G(B'))$ is a morphism in the image of $F$.
            We extend $G$ to a functor by mapping each morphism $g\from B \to B'$ to the morphism $f\from G(B) \to G(B')$ such that $g' = F(f)$.
            This morphism $f$ exists and is unique because $F$ is full and faithful.
            We have to show that $\epsln$ is a natural transformation.
            \[
                \begin{tikzcd}
                    F(G(B)) \arrow{d}{\epsln_{B}} \arrow{r}{F(G(g))} & F(G(B')) \arrow{d}{\epsln_{B'}} \\
                    B \arrow{r}{g}                                   & B'
                \end{tikzcd}
            \]
            The diagram commutes since $F(G(g)) = g' = \epsln_{B'}^{-1} \circ g \circ \epsln_{B}$.\\
            Next we define $\eta_A$ such that $F(\eta_A) = \epsln_{F(A)}^{-1}$.
            Again this is possible because $\epsln_{F(A)}^{-1}\from F(A) \to F(G(F(A)))$ is in the image of $F$.
            \[
                \begin{tikzcd}
                    A \arrow{d}{\eta_A} \arrow{r}{f} & A' \arrow{d}{\eta_{A'}} \\
                    G(F(A)) \arrow{r}{G(F(f))}       & G(F(A'))
                \end{tikzcd}
            \]
            \begin{align*}
                F(G(F(f))) & = \epsln_{F(A')}^{-1} \circ F(f) \circ \epsln_{F(A)}\\
                           & = F(\eta_{A'}) \circ F(f) \circ F(\eta_A^{-1})\\
                           & = F(\eta_{A'} \circ f \circ \eta_A^{-1})
            \end{align*}
            We obtain $G(F(f)) = \eta_{A'} \circ f \circ \eta_A^{-1}$ because $F$ is faithful.
            This shows that $\eta$ is a natural isomorphism.
            Therefore $F$ is an equivalence.
    \end{itemize}
\end{answer}

\begin{exercise}
This exercise makes precise the idea that linear algebra can equivalently
be done with matrices or with linear maps.

Fix a field $K$.  Let $\Mt$ be the category whose objects are the natural
numbers and with
\[ \Mt(m, n) = \{ n \times m \text{ matrices over } K \}. \]
Prove that $\Mt$ is equivalent to $\FDVect_K$, the category of
finite-dimensional vector spaces over $K$.
Does your equivalence involve a \emph{canonical} functor from $\Mt$ to $\FDVect_K$, or from $\FDVect_K$ to $\Mt$?
\end{exercise}
\begin{answer}
    Define a functor $F\from \Mt \to \FDVect_K$:
    \begin{align*}
        F(n) & := K^n && n \in \NN\\
        F(A\from m \to n) & := \lambda\, x. \, A \cdot x && A \text{ is a } n \times m \text{ matrix}
    \end{align*}
    Note that $F$ is indeed a functor, since matrix multiplication is associative.
    \begin{description}
        \item[$F$ is faithful:]
            Let $A, B\from m \to n$, that is $n \times m$ matrices over $K$, such that $F(A) = F(B)$.
            By using the standard basis $(e_j)_{j = 1}^m$ in $K^m$ we get:
            \[A \cdot e_j = F(A)(e_j) = F(B)(e_j) = B \cdot e_j \quad\forall 1 \le j \le m\]
            So each column of $A$ and $B$ are the same.
            Thus $A = B$.
        \item[$F$ is full:]
            Let $f\from K^m \to K^n$ be a linear function.
            We define the $n \times m$ matrix $A$ columnwise.
            The $j$-th column of $A$ is $f(e_j)$, where $e_j$ is the $j$-th standard basis vector.
            By this definition $F(A)$ and $f$ conincide on the standard basis.
            Linear algebra tells us that in this case $F(A)$ and $f$ coincide on all vectors.
            So $A$ is mapped to $f$ under $F$.
        \item[$F$ is essentially surjective on objects:]
            Let $V$ be a finite-dimensional $K$-vectorspace.
            Then $V$ has some dimension $n \in \NN$.
            Thus $V \iso K^n = F(n)$.
    \end{description}
    Thus $F$ is an equivalence and so $\Mt \eqv \FDVect_K$.
    The functor $F$ is even canonical since we didn't make arbitrary choices.
    But the functor from $\FDVect_K$ to $\Mt$ involves choosing a basis like in the case for showing that $F$ is full.
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: