### 2017-06-14 sheets/talk-7/sheet-7.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\usepackage{upgreek}

\begin{document}

\uebunghead{Talk 7}{General Limits}
\author{Leonhard Staut}

\section{Limits}

\begin{exercise}[Limit definition]
  \begin{enumerate}
  \item Verify that terminal objects are a valid instance of the general limit definition.
  \item Verify that equalizers are a valid instance of the general limit definition.
  \end{enumerate}
\end{exercise}

\begin{exercise}[Pullback monic]
  Let $\mathscr{C}$ be any category and $X, Y$ objects.
  Show that a map $X \overset{f}{\to} Y$ is monic iff the following square is a pullback.
  \[
    \begin{tikzcd}
      X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
      X \arrow{r}{f}                & Y                  
    \end{tikzcd}
  \]
\end{exercise}

\begin{exercise}[Contained limit]
  Let $D$ be a diagram.
  Find an example of a limit $L$ that is completely contained in $D$,
  meaning the vertex of $L$ is in the image of $D$.
\end{exercise}

\begin{exercise}[Cone category]
  Let $\mathscr{C}$ be any category. Let \texttt{Cone} be the category of cones
  in $\mathscr{C}$ over a fixed diagram.
  What are the limits of $\mathscr{C}$ in \texttt{Cone}? What are the colimits?
\end{exercise}

\begin{definition}
  {Constant Functor}
  For any two categories $\mathscr{C}$ and $\mathscr{D}$,
  the constant functor $\Updelta _ V : \mathscr{C} \to \mathscr{D} $ for some object $V \in \mathscr{D}$
  is defined by:
  \begin{itemize}
    \item $\forall A \in \mathscr{C}. \ \Updelta _ V (A) = V$
    \item $\forall A \overset{f}{\to} B \in \mathscr{C}. \ \Updelta _ V (f) = 1_V$
  \end{itemize}

\end{definition}

\begin{exercise}[Cones as natural transformations]
  Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
  Verify that a natural transformation $\pi : \Updelta_v \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
\end{exercise}

\begin{exercise}[Uniqueness of limits]
  Let $\mathscr{C}$ be any category. Show that limits in $\mathscr{C}$ are unique up to isomorphism.
\end{exercise}

\begin{exercise}[Limit construction]
  Recall the construction:
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K)
    \end{tikzcd}
  \]
  where $E$ is the equalizer of $s$ and $t$.
  We have shown that $E$ is a cone for any diagram $D$. Show that $E$ is also the limit,
  i.e. a universal cone over $D$.
\end{exercise}
\begin{hint}
  \begin{itemize}
  \item You may use the fact that $E$ is a cone iff $E$ is an equalizer of $s$ and $t$.
  \item You do not need the concrete definitions of $s$ and $t$.
  \end{itemize}
\end{hint}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: