### 2017-06-21 sheets/talk-7/sheet-7.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\usepackage{upgreek}

\begin{document}

\uebunghead{Talk 7}{General Limits}
\author{Leonhard Staut}

\begin{exercise}[Limit definition]
  \begin{enumerate}
  \item Verify that terminal objects are a valid instance of the general limit definition.
  \item Verify that equalizers are a valid instance of the general limit definition.
  \end{enumerate}
\end{exercise}
\begin{answer}
  Recall that general limits are universal cones over a diagram.
  Therefore, instances of general limits are defined by the shape of their diagram.
  In the following let $\mathscr{C}$ be any category.
  \begin{enumerate}
  \item
    We choose the empty category as the shape $\bf{I}$.
    Let $D : \bf{I} \to \mathscr{C}$ a diagram. A cone over $D$ is an object $U$ with
    maps to every object in the diagram. Since the diagram is empty, the cone is just
    the object $U$.
    $U$ is universal if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    Any other cone $V$ is also just an object in $\mathscr{C}$.
    Therefore $U$ is a terminal object.
  \item
    As the shape $\bf{I}$ we choose the following diagram.
    \[
      \begin{tikzcd}
        I \arrow[r,shift left, "m"] \arrow[r,shift right,swap,"n"] & J
      \end{tikzcd}
    \]
    Let $D : \bf{I} \to \mathscr{C}$ a diagram.
    A cone over $D$ is an object $U$ with maps to every object in the diagram, such that all triangles commute.
    \[
      \begin{tikzcd}
        U \arrow{d}{f_I} \arrow{dr}{f_J} & \\
        D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    Specifically, the commuting condition says that $Dm \circ f_I = f_J = Dn \circ f_I$.
    Since $f_J$ is already completely determined by $f_I$, we can ignore the arrow
    and state equivalently that the following diagram commutes.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J)
      \end{tikzcd}
    \]
    The arrow $f_J$ is simply left implicit.\\
    $U$ is a universal cone, i.e. a limit, if for any other cone $V$ in $\mathscr{C}$, there is a unique map
    $h : V \to U$.
    \[
      \begin{tikzcd}
        U \arrow{r}{f_I} & D(I) \arrow[r,shift left, "Dm"] \arrow[r,shift right,swap,"Dn"] & D(J) \\
        V \arrow{ur}{g_I} \arrow[dashed]{u}{h}& &
      \end{tikzcd}
    \]
    This is exactly the definition of an equalizer.
  \end{enumerate}
\end{answer}

\begin{exercise}[Pullback monic]
  Let $\mathscr{C}$ be any category and $X, Y$ objects.
  Show that a map $X \overset{f}{\to} Y$ is monic iff the following square is a pullback.
  \[
    \begin{tikzcd}
      X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
      X \arrow{r}{f}                & Y                  
    \end{tikzcd}
  \]
\end{exercise}

\begin{answer}
  $"\Leftarrow":$ \\
  Let $g_1, g_2 : Z \to X$ be two maps. If $f \circ g_1 = f \circ g_2$
  we have the following commuting diagram.
  \[
    \begin{tikzcd}
      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2}&   &\\
             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
             & X \arrow{r}{f}                & Y
    \end{tikzcd}
  \]
  Therefore $Z$ is a cone. Since is $X$ is a pullback we have a unique map $h : Z \to X$,
  such that the following diagram commutes.
  \[
    \begin{tikzcd}
      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2} \arrow[dashed]{dr}{h}&   &\\
             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
             & X \arrow{r}{f}                & Y
    \end{tikzcd}
  \]
  It follows that $1 \circ h = h = g_1 = g_2$.
  Because there is only one map $Z \to X$, $f$ is trivially monic.\\
  $"\Rightarrow":$ \\
  The following square certainly commutes because $f \circ 1 = f \circ 1$.
  \[
    \begin{tikzcd}
      X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
      X \arrow{r}{f}                & Y
    \end{tikzcd}
  \]
  Therefore it is a cone and we only need to show its universality.
  We have to show that, assuming f is monic, for any other cone $Z$ there is a unique map
  $h : Z \to X$ making the following diagram commute.
  \[
    \begin{tikzcd}
      Z \arrow[bend right=30]{ddr}{g_1} \arrow[bend left=30]{drr}{g_2} \arrow[dashed]{dr}{h}&   &\\
             & X \arrow{r}{1}  \arrow{d}{1}  & X \arrow{d}{f}    \\
             & X \arrow{r}{f}                & Y
    \end{tikzcd}
  \]
  We know the outer diagram commutes because $Z$ is a cone, so $f \circ g_1 = f \circ g_2$
  Therefore we have $g_1 = g_2$ because $f$ is monic.
  Since all the maps $Z \to X$
  are the same there is exactly one map which is unique.
\end{answer}

\begin{definition}
  {Cone Category}
  Let $\mathscr{C}$ be any category, $D : \bf{I} \to \mathscr{C}$ a diagram.
  \texttt{Cone(D)} is the category of cones in $\mathscr{C}$ over $D$.
  \begin{itemize}
  \item The objects of \texttt{Cone(D)} are cones in $\mathscr{C}$ over $D$, that is, if
    $(U \overset{f_I}{\to} D(I))_{I \in \bf{I}}$ is a cone in $\mathscr{C}$, then there is a corresponding object
    in \texttt{Cone(D)}.
  \item The morphisms of \texttt{Cone(D)} are morphisms between two cones in $\mathscr{C}$, that is,
    if $(U \overset{f_I}{\to} D(I))_{I \in \bf{I}}, (V \overset{g_I}{\to}~ D(I))_{I \in \bf{I}}$
    are two cones in $\mathscr{C}$ and there is a morphism $V \overset{m}{\to} U$, then there
    is a corresponding morphism in \texttt{Cone(D)}.
  \end{itemize}
\end{definition}

\begin{exercise}[Cone category]
  Consider an object $L \in$ \texttt{Cone(D)} that corresponds to a limit in $\mathscr{C}$ over $D$.
  What kind of object is $L$?\\
  What about an object $L' \in$ \texttt{Cone(D)} that corresponds to a colimit in $\mathscr{C}$ over $D$?
\end{exercise}
\begin{answer}
  If $L$ corresponds to a limit, then by definition any other cone with vertex $V$ in $\mathscr{C}$ has a unique map
  from $V$ to the limit that corresponds to a map in \texttt{Cone(D)}.
  Because the cones in $\mathscr{C}$ are exactly the objects in \texttt{Cone(D)},
  $L$ has a unique arrow from every object \texttt{Cone(D)}.
  Therefore $L$ is a terminal object.\\
  By duality $L'$ is an initial object.
\end{answer}

\begin{definition}
  {Constant Functor}
  For any two categories $\mathscr{C}$ and $\mathscr{D}$,
  the constant functor $\Updelta _ V : \mathscr{C} \to \mathscr{D} $ for some object $V \in \mathscr{D}$
  is defined by:
  \begin{itemize}
    \item $\forall A \in \mathscr{C}. \ \Updelta _ V (A) = V$
    \item $\forall A \overset{f}{\to} B \in \mathscr{C}. \ \Updelta _ V (f) = 1_V$
  \end{itemize}

\end{definition}

\begin{exercise}[Cones as natural transformations]
  Let $\mathscr{C}$ and $\bf{I}$ be any categories, $D : \bf{I} \to \mathscr{C}$ a diagram.
  Verify that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ is a cone over $D$ with vertex $V$.
\end{exercise}
\begin{answer}
  Recall that a natural transformation $\pi : \Updelta_V \to D(\bf{I})$ has components for every object $I \in \bf{I}$
  and for every morphism $m : I \to J \in \bf{I}$ the following square commutes.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
  \]
  Now we unfold the definition of $\Updelta_V$.
  \[
    \begin{tikzcd}
      \Updelta_V (I) \arrow{r}{\pi_I}  \arrow{d}{\Updelta_V (m)}  & D(I) \arrow{d}{D(m)}    \\
      \Updelta_V (J) \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I}  \arrow{d}{1_V}  & D(I) \arrow{d}{D(m)}    \\
      V \arrow{r}{\pi_J}                & D(J)
    \end{tikzcd}
    \ = \
    \begin{tikzcd}
      V \arrow{r}{\pi_I} \arrow{dr}{\pi_J} & D(I) \arrow{d}{D(m)}    \\
                     & D(J)
    \end{tikzcd}
  \]
  To summarize, the natural transformation $\pi$ has maps $\pi_I$ for all $I \in \bf{I}$,
  and the above triangle commutes for all maps $m : I \to J$.
  This is exactly the definition of a cone with vertex $V$.
  In fact this yields an alternative equivalent definition of cones.
\end{answer}

\begin{exercise}[Uniqueness of limits]
  Let $\mathscr{C}$ be any category. Show that limits in $\mathscr{C}$ are unique up to isomorphism.
\end{exercise}
\begin{answer}
  Let $U, V$ be two limits. Then there are unique morphisms $f : U \to V$ and $g : V \to U$.
  Consider the morphism $g \circ f : U \to U$. As there is a unique morphism $U \to U$, it has to be
  the identity $g \circ f = 1_U$.
  Similarly $f \circ g = 1_V$, so $U$ and $V$ are isomorphic.
\end{answer}

\begin{exercise}[Limit construction]
  Recall the construction:
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K)
    \end{tikzcd}
  \]
  where $E$ is the equalizer of $s$ and $t$.
  We have shown that $E$ is a cone for any diagram $D$. Show that $E$ is also the limit,
  i.e. a universal cone over $D$.
\end{exercise}
\begin{hint}
  \begin{itemize}
  \item You may use the fact that $E$ is a cone iff $E$ is an equalizer of $s$ and $t$.
  \item You do not need the concrete definitions of $s$ and $t$.
  \end{itemize}
\end{hint}
\begin{answer}
  To show that $E$ is a universal cone, we need to show that for any other cone $F$, there is a unique 
  morphism $h : F \to E$.
  If $F$ is a cone, then it is also an equalizer of $s$ and $t$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} & &
    \end{tikzcd}
  \]
  Since $E$ is an equalizer, and $F$ is a cone over the same diagram, we have a unique map $h : F \to E$.
  \[
    \begin{tikzcd}
      E \arrow{r}{e} &
      \displaystyle \prod_{I \in \bf{I}} D(I) \arrow[r,shift left, "s"] \arrow[r,shift right,swap,"t"] &
      \displaystyle \prod_{J \overset{q}{\to} K \in \bf{I}} D(K) \\
      F \arrow{ur}{f} \arrow[dashed]{u}{h} & &
    \end{tikzcd}
  \]

\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: