### 2017-07-02 slides/slides-10/slides-10.tex

\documentclass{beamer}

\usepackage[utf8]{inputenc}

\beamertemplatenavigationsymbolsempty
\setbeamertemplate{footline}{
	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
		\color{gray}\scriptsize\insertframenumber}}}

\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{cd, shapes, tikzmark}
\def\pathToRoot{../../}\input{\pathToRoot headers/sharedHeader}

\title{Presheaves, Representables and\\the Yoneda Lemma}
\author{Dominik Wagner}
%\date{June 7, 2017}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Contravariant Representable Functors}

  Let $\cat A$ be a locally small category and $A\in \cat A$. The \emph{contravariant representable functor} of $A$
  \begin{align*}
    H_A=\cat A(-,A)\from \cat A^\op\to \Set
  \end{align*}
  is defined by
  \begin{itemize}
  \item $H_A(B)=\cat A(B,A)$ for objects $B\in\cat A$ and
  \item for maps $B'\xrightarrow{f} B$:
    \begin{align*}
      H_A(f)=-\of f\from \cat A(B,A)&\to \cat A(B',A)\\
                                      g &\mapsto g\of f\\\\
      \xymatrix{B' \ar[r]^{f} & B \ar[r]^{g} & A }
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Yoneda Embedding}
  Let $\cat A$ be locally small. The \emph{Yoneda embedding} of $\cat A$ is the functor
  \begin{align*}
    H_-\from \cat A\to[\cat A^\op,\Set]
  \end{align*}
  defined by
  \begin{itemize}
  \item $H_-(A)=H_A$ for objects $A\in\cat A$ and
  \item for maps $A\xrightarrow{f}A'$, $H_-(f)$ is the natural transformation defined for $B\in \cat A$ by
    \begin{align*}
      H_A(B)=\cat A(B,A)&\to H_{A'}(B)=\cat A(B,A')\\
      g&\mapsto  f\of g\\
    \end{align*}
  \end{itemize}
  \vspace{-0.8cm}
  \includegraphics{nattrans.pdf}
\end{frame}

\begin{frame}
  \frametitle{Yoneda Lemma and Yoneda Principle}
  \begin{lemma}[Yoneda]
    Let $\cat A$ be locally small. Then
    \begin{align*}
      [\cat A^\op,\Set](H_A,F)\iso F(A)
    \end{align*}
    naturally in $A\in\cat A$ and $F\in [\cat A^\op,\Set]$.
  \end{lemma}
  \begin{theorem}
    The Yoneda embedding $H_-\from \cat A\to[\cat A^\op,\Set]$ is full and faithful.
  \end{theorem}
  \begin{corollary}[Yoneda Principle]
    Let $\cat A$ be locally small and $A,A'\in\cat A$. Then
    \begin{align*}
      H_A\iso H_{A'} \iff A\iso A'
    \end{align*}
  \end{corollary}
\end{frame}
\end{document}

% vim mode line:
% vim: ts=2 sts=2 sw=2 expandtab