### 2017-07-12 slides/talk-12/header.tex
Add slides for talk 12

\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{semantic}
\usepackage{tikzsymbols}  % This defines \Cat, that we want to redefine
\usepackage{stmaryrd}  % This defines \mapsfrom, that we want to redefine
\usepackage{tikz}
\usetikzlibrary{cd}
\usepackage{adjustbox}
\usepackage{transparent}

\beamertemplatenavigationsymbolsempty
\setbeamertemplate{footline}{
	\raisebox{0.5em}{\makebox[\paperwidth-0.5em][r]{
		\color{gray}\scriptsize\insertframenumber}}}

\input{../../headers/definitions}