### 2017-07-18 sheets/talk-12/sheet-12.tex
Add exercises for talk 12

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\graphicspath{{graphics/}}

\begin{document}
\uebunghead{Talk 12}{Monoidal categories and string diagrams}
\author{Felix Rech}

\begin{exercise}
  Let $A$, $B$, $C$, $D$, be objects in a monoidal category.
  Construct a morphism
  \[
    \roundddd{\rounddd{(A \otimes I) \otimes B} \otimes C} \otimes D
    \to A \otimes \roundddd{B \otimes \rounddd{C \otimes (I \otimes D)}}.
  \]
  Can you find another?
\end{exercise}

\begin{exercise}
  Convert the following algebraic equations into graphical language.
  Which would you expect to be true in any symmetric monoidal category?
  \begin{enumerate}
    \item
      $(g \otimes 1) \of \sigma \of (f \otimes 1)
        = (f \otimes 1) \of \sigma \of (g \otimes 1)$ \\
      for $A \xrightarrow{f, g} A$.

    \item
      $h \of (1 \otimes \lambda) \of (1 \otimes (f \otimes 1)) \of (1 \otimes \lambda^{-1}) \of g
        = h \of g \of \lambda \of (f \otimes 1) \of \lambda^{-1}$ \\
      for $A \xrightarrow{g} B \otimes C$, $I \xrightarrow{f} I$ and $B \otimes C \xrightarrow{h} D$.

    \item
      $\rho \of (1 \otimes f) \of \alpha \of (\sigma \otimes 1)
        = \lambda \of (f \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha$ \\
      for $A \otimes B \xrightarrow{f} f$.
  \end{enumerate}
\end{exercise}

\begin{exercise}
    Consider the following diagrams in the graphical calculus:

    \centering \includegraphics{monoidal-braided-symmetric-comparison}

    \begin{enumerate}
      \item Which diagrams are equal as morphisms in a monoidal category?
      \item Which diagrams are equal as morphisms in a braided monoidal category?
      \item Which diagrams are equal as morphisms in a symmetric monoidal category?
    \end{enumerate}
\end{exercise}

\pagebreak
\begin{exercise}
  Verify the correctness of the following graphical equations by
  \begin{enumerate}
    \item the correctness theorem of the graphical calculus,
    \item substitution in string diagrams using the axioms of braided monoidal categories,
    \item construction of a commuting diagram from the axioms.
  \end{enumerate}
  \begin{enumerate}[label=(\roman*)]
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_1}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_2}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_3}}}$
      ($\sigma_{A, I} = \lambda_{A}^{-1} \of \rho_A$)
  \end{enumerate}
\end{exercise}

\begin{exercise}
  Show that $\Set$ is a symmetric monoidal category under $I \coloneqq \emptyset$ and $A \otimes B \coloneqq A + B + (A \times B)$, where we write $\times$ for Cartesian product of sets and $+$ for disjoint union of sets.
\end{exercise}

\begin{exercise}
  Assume that we have morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon, \varepsilon'} I$ in a monoidal category such that $\eta$ and $\varepsilon$ satisfy one snake equation and $\eta$ and $\varepsilon'$ satisfy the other snake equation.
  Show that this already implies $\varepsilon = \varepsilon'$.
\end{exercise}

\begin{exercise}
  In a monoidal category, show that
  \begin{enumerate}
    \item if an initial object $0$ exists and $L \dashv R$, then $L \otimes 0 \iso 0 \iso 0 \otimes R$;
    \item if a terminal object $1$ exists and $L \dashv R$, then $R \otimes 1 \iso 1 \iso 1 \otimes L$.
  \end{enumerate}
\end{exercise}

\begin{definition}{Idempotent, splitting}
  An endomorphism $A \xrightarrow{f} A$ is called \demph{idempotent} when $f \of f = f$. An idempotent $A \xrightarrow{f} A$ \demph{splits} when there exists an object $B$ and morphisms $A \xrightarrow{g} B$ and $B \xrightarrow{h} A$ such that $f = h \of g$ and $g \of h = 1$.
\end{definition}

\begin{exercise}
  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily $R$).
\end{exercise}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: