### 2017-07-19 sheets/talk-11/sheet-11.tex

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}
\usepackage{semantic}

\begin{document}

% Use Basis x or Talk x, where x is the number of the session
\uebunghead{Talk 11}{Dependent Types}
\author{Nikita Ziuzin}

\begin{hint}
  Read sections $2.1.3, 2.1.5, 2.1.6$ in Martin Hofmann's \href{http://www.irif.fr/~mellies/mpri/mpri-ens/articles/hofmann-syntax-and-semantics-of-dependent-types.pdf}{Syntax and Semantics of Dependent Types} and section $1$ in Simon Huber's \href{http://www.cse.chalmers.se/~simonhu/misc/lic.pdf}{A Model of Type Theory in Cubical Sets}.
\end{hint}

\begin{exercise}
  Give the sets $\mathit{Ty}_F(\Gamma)$ and $\mathit{Ter}_F(\Gamma, A)$ for
  the presheaf model.
\end{exercise}
\begin{answer}
  \[
    \mathit{Ty}_F(\Gamma) := \mathit{Psh(\Gamma)}
  \]
  \[
    \mathit{Ter}_F(\Gamma, A) := \{(a\rho \in A\rho)_{I \in \cat{C}, \rho \in \Gamma(I)} | \forall f \from J \to I, a\rho f = a(\rho f) \}
  \]
\end{answer}

\begin{exercise}
  Let $(\cat C, \mathcal F)$ be a category with families.  Check that the
  following square is a pullback in $\cat{C}$:
  \[
    \begin{tikzcd}
      \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
      \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  where $\Delta.A \vdash q : A p_\Delta$
\end{exercise}
\begin{answer}
  Let $\Delta'$ be a context and $p' \from \Delta' \to \Delta$ and $\sigma'
  \from \Delta' \to \Gamma.A$ be two morphisms such that $\sigma \circ p' =
  p_\Gamma \circ \sigma'$. Therefore the diagram
  \[
    \begin{tikzcd}
      \Delta' \arrow[bend right=30]{ddr}{p'} \arrow[dashed]{dr}{\delta} \arrow[bend left=30]{drr}{\sigma'}&   &\\
          & \Delta.A\sigma \arrow{r}{(\sigma \circ p_\Delta, q)} \arrow{d}{p_\Delta} & \Gamma.A \arrow{d}{p_\Gamma}\\
          & \Delta \arrow{r}{\sigma} & \Gamma
    \end{tikzcd}
  \]
  (without $\delta$) is commutative. We must show that there exists a unique
  $\delta \from \Delta' \to \Delta.A\sigma$ such that $(\sigma \circ p_\Delta,
  q) \circ \delta = \sigma'$ and $p_\Delta \circ \delta = p'$.

  We can decompose $\sigma'$ as $\sigma' = (p_\Gamma, q_\Gamma) \circ \sigma' =
  (p_\Gamma \circ \sigma', q_\Gamma\sigma') = (\sigma \circ p', M)$, where $M
  \equiv q_\Gamma\sigma'$ (of type $\Delta' \vdash A(\sigma \circ p')
  ~\text{type}$). Let $\delta$ be the morphism
  \[
    \delta \equiv (p', M) \from \Delta' \to \Delta.A\sigma
  \]
  which is suitable since $p_\Delta \circ \delta = p'$ and
  \[
    (\sigma \circ p_\Delta, q) \circ \delta =
    (\sigma \circ p_\Delta \circ \delta, q\delta) =
    (\sigma \circ p', q_\Gamma\sigma') =
    (p_\Gamma \circ \sigma', q_\Gamma\sigma') =
    (p_\Gamma, q_\Gamma) \circ \sigma' =
    \sigma'
  \]

  For uniqueness assume morphism $\delta' \from \Delta' \to \Delta.A\sigma$
  such that $(\sigma \circ p_\Delta, q) \circ \delta' = \sigma'$ and $p_\Delta
  \circ \sigma' = p'$. As before we have $\delta' = (p_\Delta \circ \delta',
  q\delta') = (p', N)$ where $N \equiv q\delta'$. The following equalities
  hold:
  \[
    N = q\delta' = q_\Gamma(\sigma \circ p_\Delta \circ \delta', q\delta') = q_\Gamma((\sigma \circ p_\Delta, q) \circ \delta') = q_\Gamma\sigma' = M
  \]
  and therefore $\delta' = \delta$.
\end{answer}

\begin{exercise}
  Recall the notation $\int_\cat{C}\Gamma$ for the category of elements of a
  presheaf $\Gamma$ (section $1.2$ p.$16$ in Huber's thesis).  Check that
  substitution in $A\from (\int_\cat{C}\Gamma)^\op \to \Set$ with $\sigma \from
  \Delta \to \Gamma$ corresponds to precomposing with $\int_\cat{C} \sigma :
  \int_\cat{C} \Delta \to \int_\cat{C} \Gamma$ induced by $\sigma$.

  Then verify that this construction imposes a functor $\int_\cat{C} \from
  \mathit{Psh}(\cat{C}) \to \CAT$.
\end{exercise}
\begin{answer}
  Indeed, the operation of $\sigma$ on such objects is a precomposition
  satisfying the naturality axiom:
  \[
    \int_\cat{C}\sigma(I, \rho) := (I, \sigma_I \rho)
  \]
  where $\rho \in \Gamma(I), I \in \cat{C}$

  Let's check that $\int_\cat{C}$ takes us into a category. Recall the
  definition of $\int_\cat{C}\Gamma$:
  \begin{itemize}
    \item objects are pairs $(I, \rho)$ where $I \in \cat{C}$ and $\rho \in \Gamma(I)$
    \item a morphism $(J, \rho') \to (I, \rho)$ is a morphism $f \from J \to I$
      in $\cat{C}$ such that $\rho' = \rho f$.
  \end{itemize}

  We know our objects and morphisms, we get identities and composition by
  the equations $\rho 1 = \rho$, $(\rho f) g = \rho (f g)$ and the fact that
  morphisms $f$ are valid morphisms in the category $\cat{C}$.

  Let's now see that $\int_\cat{C}\sigma$ is a functor in $\CAT$. We have a
  mapping for objects $\int_\cat{C}\sigma(I, \rho)$ given above. The operation
  on arrows is as follows, where $f \from J \to I$, $g \from K \to J$:
  \[
    \int_\cat{C}\sigma(f) := \sigma_{I \to J} \Gamma_f
  \]
  Then, keeping in mind that composition $fg$ is applied on the right we get
  $\int_\cat{C}\sigma(fg) = \sigma_{I \to K} \Gamma_{fg} = \sigma_{I \to K}
  \lambda \rho.\rho (fg) = \sigma_{I \to K} \lambda \rho. (\rho f)g = \sigma_{J
  \to K} (\sigma_{I \to J} \lambda \rho.\rho f) g = (\sigma_{J \to K} \lambda
  \rho.\rho g) (\sigma_{I \to J} \lambda.\rho f) = (\sigma_{J \to K} \Gamma_g)
  (\sigma_{I \to J} \Gamma_f)$. Which, together with $\rho 1 = \rho$ confirms
  that $\int_\cat{C}\sigma$ is indeed a functor.

  Thus, we conclude that $\int_\cat{C} \from \mathit{Psh}(\cat{C}) \to \CAT$ is
  a functor.
\end{answer}

\begin{exercise}
  From the slides we know that if $\Gamma \vdash A, B ~\text{type}$, i.e. $B$
  does not depend on $A$, then $\Pi x: A. B$ corresponds to the ordinary
  non-dependent function type:
  \begin{equation*}
    A \to B \vcentcolon= \Pi x: A. B
  \end{equation*}

  Give a term of type $A \to A$, and the elements in $\mathit{Ty}(\Gamma)$ and
  $\mathit{Ter}(\Gamma, A \to A)$ corresponding to the type and the term.

  \emph{Hint}: Read section $1.2.1$ in Huber's thesis for the general case.
\end{exercise}
\begin{answer}
  For a CwF $(\cat{C}, F), I, J \in ob(\cat{C}), f \from J \to I$ we give the
  element of $\mathit{Ty}(\Gamma)$ corresponding to the type $A \to A$:
  \[
    (A \to A)\rho = \{w | w_f \in A(\rho f), u \in A(\rho f)\} = \{\lambda I f u.u\}
  \]
  The required equalities hold trivially for any $g \from K \to J$, $((\lambda
  I f u.u)_f u) g = (\lambda I f u.u)_{fg} (ug)$.

  $\Gamma, x : A \vdash x : A$ corresponds to $q$ in the model. Thus $\Gamma
  \vdash \lambda x.x : A -> A$ corresponds to $\lambda q$ being defined as
  follows:
  \[
    ((\lambda q) \rho)_f u = q(\rho f, u) = u
  \]
\end{answer}

\begin{exercise}
  Suppose we have two types $\Gamma \vdash A,B \text{ type}$ each not depending
  on the other.  Then we can define a product type in the presheaf model in the
  following way:
  \[
    (A \times B)\rho := \{(a,b) | a \in A\rho \land b \in B\rho\}
  \]

  Now suppose that we have dependent types $\Gamma \vdash C \text{ type}$ and
  $\Gamma, C \vdash D \text{ type}$.

  Define the product type that involves a dependent type (i.e. the dependent
  sum type) in the presheaf model.

  \[
    (\Sigma AB)\rho := ???
  \]

  \emph{Hint}: The definition is \emph{almost} identical to the one given
  above. But what do you need to change? To recall the presheaf model, read
  section $1.2$ in Huber's thesis.
\end{exercise}
\begin{answer}
  \[
    (\Sigma AB)\rho := \{(a,b) | a \in A\rho \land b \in B(\rho, a)\}
  \]
\end{answer}

\begin{exercise}
  Read Hofmann's section on the type of natural numbers (section 2.1.3) and
  familiarize yourself with identity types (ibid., section 2.1.5)

  By analogy to the type of natural numbers define the rules for a list type
  former which to any type $A$ associates a type $\mathit{List}(A)$ consisting
  of finite sequences of elements of $A$.

  \emph{Hint}: think of lists as inductively generated from the empty list by
  successive additions of elements of $A$ ("cons").

  Define a length function of type $\mathit{List}(A) \to \mathbb{N}$ and define
  a type $\mathit{Vec}_A(n)$ of lists of length $n$ for each $n : \mathbb{N}$
  using lists, the identity type and the $\Sigma$-type.
\end{exercise}
\begin{answer}
  First let's define the type formers:
  \[
    \inference[List-F]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathbf{List_A} ~\text{type}}\quad
    \inference[List-I-nil]{\Gamma \vdash A ~\text{type}}{\Gamma \vdash \mathit{nil_A} : \mathbf{List_A}}\quad
    \inference[List-I-cons]{\Gamma \vdash a : A \quad \Gamma \vdash x : \mathbf{List_A}}{\Gamma \vdash \mathit{Cons(a, x)} : \mathbf{List_A}}
  \]

  \[
    \inference[List-E]{
      \Gamma, l: \mathbf{List_A} \vdash B ~\text{type}\\
      \Gamma \vdash H_{nil} : B[nil_A/l]\\
      \Gamma, a : A, l : \mathbf{List_A}, x: B \vdash H_{cons} : B[Cons(a, l)/l]\\
      \Gamma \vdash M : \mathbf{List_A}
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, M) : B[M/l]}
  \]

  \[
    \inference[List-C-nil]{
      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, nil) : B[nil/l]
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B]H_{cons}, nil) = H_{nil} : B[nil/l]}
  \]

  \[
    \inference[List-C-cons]{
      \Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(N, M)) : B[Cons(N, M)/l]
    }{\Gamma \vdash R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x: B] H_{cons}, Cons(N, M)) = D}
  \]
  where $D = H_{cons}[N/a, M/l, R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]B}(H_{nil}, [a : A, l: \mathbf{List_A}, x : B] H_{cons}, M)/x] : B[Cons(N, M)/l]$.

  Now we can define the length function:
  \[
    \mathit{Length(l) \vcentcolon= R^{\mathbf{List_A}}_{[l: \mathbf{List_A}]\mathbb{N}}(0, [a : A, l: \mathbf{List_A}, x : \mathbb{N}] Suc(x), l) : \mathbb{N}}
  \]

  And finally the $\mathit{Vec_A(n)}$ type:
  \[
    \mathit{Vec_A(n)} \vcentcolon= \Sigma l: \mathbf{List_A}. Id_\mathbb{N}(n, \mathit{Length}(l))
  \]

  Note that it is not the only way to define $\mathit{Vec_A(n)}$, as one could
  define it with the type formers analogously to our type $\mathit{List_A}$.
\end{answer}

\begin{exercise}
  Read the section on universes in Hofmann's notes (2.1.6) as well as the
  material for the previous exercise.

  Give the rules for universe $U$ containing a code for $\hat{0}$ for the empty
  type and a code $\hat{1}$ for the unit type $1$. Show that in a type theory
  which supports natural numbers, this universe, and the empty type itself the
  following type in the empty context is inhabited
  \[
    \diamond \vdash Id_\mathbb{N}(0, Suc(0)) \to 0 \text{ type}
  \]
  corresponding to Peano's fourth axiom $1 \neq 0$.

  \emph{Hint}: define using $R^\mathbb{N}$ a function $f \from \mathbb{N} \to
  U$ such that $\diamond \vdash f 0 = \hat{1}: U$ and $\diamond \vdash
  f(Suc(0)) = \hat{0} : U$
\end{exercise}
\begin{answer}
  \[
    \inference[U-0]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{0} : U}\quad
    \inference[U-0-Ty]{\Gamma \vdash \hat{0}: U}{\Gamma \vdash El(\hat{0}) = 0~\text{type}}
  \]

  \[
    \inference[U-1]{\vdash \Gamma~\text{ctxt}}{\Gamma \vdash \hat{1} : U}\quad
    \inference[U-1-T]{\Gamma \vdash \hat{1} : U}{\Gamma \vdash El(\hat{1}) = 1~\text{type}}
  \]

  \[
    f(n) = R^\mathbb{N}_{[n:\mathbb{N}]U}(\hat{1}, [n: \mathbb{N}, x: U]\hat{0}, n)
  \]

  Now to get an inhabitant of that type let's use $\mathit{Subst}$:
  \[
    P : Id(0, Suc(0)) \vdash \mathit{Subst}(P, El(f(0))) : El(f(Suc(0)))
  \]
  Since the inhabitant to $El(f(0)) = El(\hat{1}) = 1~\text{type}$, namely
  $\star$, is always present by the formation rules, and $El(f(Suc(0))) =
  0~\text{type}$ we obtain an inhabitant to the required type.
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: