### 2017-07-26 sheets/talk-12/sheet-12.tex
Add example solutions for talk 12

\def\pathToRoot{../../}\input{\pathToRoot headers/uebungHeader}

\usetikzlibrary{fit}
\usetikzlibrary{knots}
\usetikzlibrary{hobby}
\tikzstyle{morphism}=[draw, minimum size=0.5cm, semithick]
\tikzstyle{phantom morphism}=[minimum size=0.5cm]
\tikzstyle{object}=[semithick]
\tikzstyle{cap object}=[object, out=90, in=90, looseness=2]
\tikzstyle{cup object}=[object, out=270, in=270, looseness=2]
\tikzstyle{highlight box}=[draw, dotted]

\graphicspath{{graphics/}}

\begin{document}
\uebunghead{Talk 12}{Monoidal categories and string diagrams}
\author{Felix Rech}

\begin{exercise}
  Let $A$, $B$, $C$, $D$, be objects in a monoidal category.
  Construct a morphism
  \[
    \roundddd{\rounddd{(A \otimes I) \otimes B} \otimes C} \otimes D
    \to A \otimes \roundddd{B \otimes \rounddd{C \otimes (I \otimes D)}}.
  \]
  Can you find another?
\end{exercise}

\begin{answer}
  We construct the morphism as
  \[ \begin{tikzcd}
    (((A \otimes I) \otimes B) \otimes C) \otimes D
      \arrow[d, "((\rho_A \otimes 1_B) \otimes 1_C) \otimes 1_D"] \\
    ((A \otimes B) \otimes C) \otimes D
      \arrow[d, "\alpha_{A, B, C} \otimes 1_D"] \\
    (A \otimes (B \otimes C)) \otimes D
      \arrow[d, "\alpha_{A, B \otimes C, D}"] \\
    A \otimes ((B \otimes C) \otimes D)
      \arrow[d, "1_A \otimes \alpha_{B, C, D}"] \\
    A \otimes (B \otimes (C \otimes D))
      \arrow[d, "1_A \otimes (1_B \otimes (1_C \otimes \lambda_D^{-1}))"] \\
    A \otimes (B \otimes (C \otimes (I \otimes D))).
  \end{tikzcd} \]
  We have no more information about the category except that it is monoidal.
  In this setting we cannot construct two different morphisms of the same type, which follows from the coherence theorem.
\end{answer}

\begin{exercise}
  Convert the following algebraic equations into graphical language.
  Which would you expect to be true in any symmetric monoidal category?
  \begin{enumerate}
    \item
      $(g \otimes 1) \of \sigma \of (f \otimes 1)
        = (f \otimes 1) \of \sigma \of (g \otimes 1)$ \\
      for $A \xrightarrow{f, g} A$.

    \item
      $h \of (1 \otimes \lambda) \of (1 \otimes (f \otimes 1)) \of (1 \otimes \lambda^{-1}) \of g
        = h \of g \of \lambda \of (f \otimes 1) \of \lambda^{-1}$ \\
      for $A \xrightarrow{g} B \otimes C$, $I \xrightarrow{f} I$ and $B \otimes C \xrightarrow{h} D$.

    \item
      $\rho \of (1 \otimes f) \of \alpha \of (\sigma \otimes 1)
        = \lambda \of (f \otimes 1) \of \alpha^{-1} \of (1 \otimes \sigma) \of \alpha$ \\
      for $A \otimes B \xrightarrow{f} I$.
  \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (in_f) at (0, 0);
      \node[morphism] (f) at (0, 1) {$f$};
      \node[phantom morphism] (f') at (1, 2.5) {};
      \coordinate (out_f) at (1, 3.5);

      \coordinate (in_g) at (1, 0);
      \node[phantom morphism] (g') at (1, 1) {};
      \node[morphism] (g) at (0, 2.5) {$g$};
      \coordinate (out_g) at (0, 3.5);

      \begin{knot}[clip width=8]
        \strand[object] (in_f)
          to node[auto] {$A$} (f.south)
          (f.north) to[out=90, in=270] (f'.south)
          to (f'.north)
          to node[auto] {$A$} (out_f);
        \strand[object] (in_g)
          to node[auto] {$A$} (g'.south)
          to (g'.north)
          to[out=90, in=270] (g.south)
          (g.north) to node[auto] {$A$} (out_g);
      \end{knot}
    \end{tikzpicture}}}
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (in_f) at (0, 0);
      \node[morphism] (f) at (0, 1) {$g$};
      \node[phantom morphism] (f') at (1, 2.5) {};
      \coordinate (out_f) at (1, 3.5);

      \coordinate (in_g) at (1, 0);
      \node[phantom morphism] (g') at (1, 1) {};
      \node[morphism] (g) at (0, 2.5) {$f$};
      \coordinate (out_g) at (0, 3.5);

      \begin{knot}[clip width=8]
        \strand[object] (in_f)
          to node[auto] {$A$} (f.south)
          (f.north) to[out=90, in=270] (f'.south)
          to (f'.north)
          to node[auto] {$A$} (out_f);
        \strand[object] (in_g)
          to node[auto] {$A$} (g'.south)
          to (g'.north)
          to[out=90, in=270] (g.south)
          (g.north) to node[auto] {$A$} (out_g);
      \end{knot}
    \end{tikzpicture}}} \]

  \item
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (in) at (0, 0);

      \node[phantom morphism] (g_left) at (-1, 1) {};
      \node[phantom morphism] (g_right) at (1, 1) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$g$};

      \node[morphism] (f) at (0, 2) {$f$};

      \node[phantom morphism] (h_left) at (-1, 3) {};
      \node[phantom morphism] (h_right) at (1, 3) {};
      \node[morphism, fit=(h_left) (h_right), inner sep=0] (h) {};
      \node at (h) {$h$};

      \coordinate (out) at (0, 4);

      \draw[object] (in) to node[auto] {$A$} (g);
      \draw[object] (-1, 0 |- g.north) to node[auto] {$B$} (-1, 0 |- h.south);
      \draw[object] (1, 0 |- g.north) to node[auto] {$C$} (1, 0 |- h.south);
      \draw[object] (h) to node[auto] {$D$} (out);
    \end{tikzpicture}}}
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (in) at (0, 0);

      \node[morphism] (f) at (-1, 1) {$f$};

      \node[phantom morphism] (g_left) at (-1, 2) {};
      \node[phantom morphism] (g_right) at (1, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$g$};

      \node[phantom morphism] (h_left) at (-1, 3) {};
      \node[phantom morphism] (h_right) at (1, 3) {};
      \node[morphism, fit=(h_left) (h_right), inner sep=0] (h) {};
      \node at (h) {$h$};

      \coordinate (out) at (0, 4);

      \draw[object] (in) to node[auto] {$A$} (g);
      \draw[object] (-1, 0 |- g.north) to node[auto] {$B$} (-1, 0 |- h.south);
      \draw[object] (1, 0 |- g.north) to node[auto] {$C$} (1, 0 |- h.south);
      \draw[object] (h) to node[auto] {$D$} (out);
    \end{tikzpicture}}} \]

  \item
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (A_in) at (0, 0);
      \coordinate (C_in) at (1, 0);
      \coordinate (B_in) at (2, 0);
      \node[phantom morphism] (C_1) at (0, 1.5) {};

      \node[phantom morphism] (f_left) at (1, 1.5) {};
      \node[phantom morphism] (f_right) at (2, 1.5) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$f$};

      \coordinate (out) at (0, 2);

      \draw[object] (B_in) to node[auto, at start] {$B$} (B_in |- f.south);

      \begin{knot}[clip width=8]
        \strand[object, out=90, in=270] (A_in)
          to node[auto, at start] {$A$} (C_in |- f.south);
        \strand[object, out=90, in=270] (C_in)
          to node[auto, at start] {$C$} (C_1.south)
          to (out);
      \end{knot}
    \end{tikzpicture}}}
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \coordinate (A_in) at (0, 0);
      \coordinate (C_in) at (1, 0);
      \coordinate (B_in) at (2, 0);
      \node[phantom morphism] (C_1) at (2, 1.5) {};

      \node[phantom morphism] (f_left) at (0, 1.5) {};
      \node[phantom morphism] (f_right) at (1, 1.5) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$f$};

      \coordinate (out) at (2, 2);

      \draw[object] (A_in) to node[auto, at start] {$A$} (A_in |- f.south);

      \begin{knot}[clip width=8]
        \strand[object, out=90, in=270] (C_in)
          to node[auto, at start] {$C$} (C_1.south)
          to (out);
        \strand[object, out=90, in=270] (B_in)
          to node[auto, at start] {$B$} (C_in |- f.south);
      \end{knot}
    \end{tikzpicture}}} \]
  \end{enumerate}

  The correctness theorem for the graphical calculus on symmetric monoidal categories tells us that two diagrams describe the same morphism if and only if they are isotopic in four-dimensional space.
  This means essentially that nodes can move accross edges and that edges can move \textit{through} each other.
  With this we can see that equations (b) and (c) hold, but not equation (a).
\end{answer}

\begin{exercise}
    Consider the following diagrams in the graphical calculus:

    \centering \includegraphics{monoidal-braided-symmetric-comparison}

    \begin{enumerate}
      \item Which diagrams are equal as morphisms in a monoidal category?
      \item Which diagrams are equal as morphisms in a braided monoidal category?
      \item Which diagrams are equal as morphisms in a symmetric monoidal category?
    \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item
    We regard the diagrams up to planar isotopy.
    Thus diagrams (1) and (2) are equal but (3) and (4) are not equal to any other diagram.
  \item
    We regard the diagrams up to spacial isotopy.
    Thus diagrams (1), (2) and (3) are equal but (4) is not equal to any other diagram.
  \item
    We regard the diagrams up to four-dimensional isotopy.
    Thus all diagrams are equal.
  \end{enumerate}
\end{answer}

\ifdefined\issolution \else \pagebreak \fi
\begin{exercise}
  Verify the correctness of the following graphical equations in braided monoidal categories by
  \begin{enumerate}
    \item the correctness theorem of the graphical calculus,
    \item substitution in string diagrams using the axioms of braided monoidal categories and the correctness theorem for the graphical calculus on general monoidal categories,
    \item construction of a commuting diagram from the axioms and the coherence theorem for general monoidal categories.
  \end{enumerate}
  \begin{enumerate}[label=(\roman*)]
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_1}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_2}}}$
    \item $\vcenter{\hbox{\includegraphics{proofs_by_hand_3}}}$
      ($\sigma_{A, I} = \lambda_{A}^{-1} \of \rho_A$)
  \end{enumerate}
\end{exercise}

\begin{answer}
  \begin{enumerate}
  \item
    In three-dimensional space, the diagrams are obviously isotopic.
    The equations follow directly from the correctness theorem.
  \item
    \begin{enumerate}
    \item
      \[
        \vcenter{ \hbox{ \begin{tikzpicture}
          \node[phantom morphism] (in) at (0, -0.5) {};
          \node[phantom morphism] (f_left) at (1, 2) {};
          \node[phantom morphism] (f_right) at (2, 2) {};
          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
          \node[phantom morphism] (out_1) at (0, 4) {};
          \node[phantom morphism] (out_2) at (1, 4) {};
          \node[phantom morphism] (out_3) at (2, 4) {};

          \begin{knot}[clip width=8]
            \strand[object, out=90, in=270] (in)
              to (in |- f.north)
              to (out_2);
            \strand[object, out=90, in=270] (f_left) to (out_1);
            \strand[object, out=90, in=270] (f_right) to (out_3);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
          \node[phantom morphism] (in) at (2.5, -0.5) {};
          \node[phantom morphism] (f_left) at (0.5, 1) {};
          \node[phantom morphism] (f_right) at (1.5, 1) {};
          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
          \node[phantom morphism] (out_1) at (1, 4) {};
          \node[phantom morphism] (out_2) at (2, 4) {};
          \node[phantom morphism] (out_3) at (3, 4) {};

          \begin{knot}[clip width=8]
            \strand[object, out=90, in=270] (in)
              to (in |- f.north)
              to (1, 2.5)
              to (out_2);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90] (f_left.north)
              .. (1.6, 1.92)
              .. (out_2 |- 0, 2.5)
              to (out_1);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
              (f_right.north)
              .. (1.9, 1.83)
              .. (out_3 |- 0, 2.5)
              to (out_3);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}
          \node[phantom morphism] (in) at (2, 0) {};
          \node[phantom morphism] (f_left) at (0, 1) {};
          \node[phantom morphism] (f_right) at (1, 1) {};
          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
          \node[phantom morphism] (out_1) at (0, 4.5) {};
          \node[phantom morphism] (out_2) at (1, 4.5) {};
          \node[phantom morphism] (out_3) at (2, 4.5) {};

          \begin{knot}[clip width=8]
            \strand[object, out=90, in=270] (in)
              to (in |- f.north)
              to (0, 3)
              to (out_2);
            \strand[object, in=-90, out=90] (f_left.north)
              to (0, 2)
              to (1, 3)
              to (out_1);
            \strand[object, in=-90, out=90]
              (f_right.north)
              to (2, 2.5)
              to (out_3);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}
          \node[phantom morphism] (in) at (2, 0) {};
          \node[phantom morphism] (f_left) at (0, 1) {};
          \node[phantom morphism] (f_right) at (1, 1) {};
          \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
          \node[phantom morphism] (out_1) at (0, 4.5) {};
          \node[phantom morphism] (out_2) at (1, 4.5) {};
          \node[phantom morphism] (out_3) at (2, 4.5) {};

          \begin{knot}[clip width=8]
            \strand[object, out=90, in=270] (in)
              to (in |- f.north)
              to (1, 2.5)
              to (out_2);
            \strand[object, out=90, in=270] (f_left) to (out_1);
            \strand[object, out=90, in=270] (f_right) to (2, 2.5) to (out_3);
          \end{knot}
        \end{tikzpicture} } }
      \]
      In step one we apply naturality of $\sigma^{-1}$, in step two we apply one hexagon equality and in step three we cancel inverses.
      Note that we implicitly use the equality from (iii) to justify that we cross the invisible unit object below the box with another object in step one.

    \item
      \[
        \vcenter{ \hbox{ \begin{tikzpicture}
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (1, 1)
              to (2, 2)
              to (2, 3);
            \strand[object, in=-90, out=90]
              (1, 0)
              to (0, 1)
              to (0, 2)
              to (1, 3);
            \strand[object, in=-90, out=90]
              (2, 0)
              to (2, 1)
              to (1, 2)
              to (0, 3);
          \end{knot}

          % \draw[highlight box] (-0.2, 0) rectangle (2.2, 2);
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (2, 2)
              to (2, 3);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
              (1, 0)
              .. (0.9, 1)
              .. (0, 2)
              to (1, 3);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
              (2, 0)
              .. (1.1, 1)
              .. (1, 2)
              to (0, 3);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[use Hobby shortcut]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (0, 1)
              to (2, 3);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
              (1, 0)
              to (2, 1)
              .. (1.1, 2)
              .. (1, 3);
            \strand[object, in angle=-90, out angle=90, in=-90, out=90]
              (2, 0)
              to (1, 1)
              .. (0.9, 2)
              .. (0, 3);
          \end{knot}

          % \draw[highlight box] (-0.2, 1) rectangle (2.2, 3);
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (0, 1)
              to (1, 2)
              to (2, 3);
            \strand[object, in=-90, out=90]
              (1, 0)
              to (2, 1)
              to (2, 2)
              to (1, 3);
            \strand[object, in=-90, out=90]
              (2, 0)
              to (1, 1)
              to (0, 2)
              to (0, 3);
          \end{knot}
        \end{tikzpicture} } }
      \]
      We apply one hexagon equality in step one, naturality of $\sigma$ in step two and the other hexagon equality in step three.

    \item
      \[
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90] (-1, 0) to (1, 4.5);
            \strand[object, dashed, in=-90, out=90] (1, 0) to (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, in=-90, out=90]
              (1, 0)
              to (-1, 1.5)
              to (1, 3)
              to (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, in=-90, out=90]
              (1, 0)
              to (-1, 1.4)
              to (-1, 1.6)
              to (1, 3.1)
              (-1, 1.4)
              to (1, 2.9)
              to (1, 3.1)
              to (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object, in=-90, out=90]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, in=-90, out=90]
              (1, 0)
              to (-1, 1)
              to (-1, 2)
              to (1, 3.5)
              (-1, 1)
              to (1, 2.5)
              to (1, 3.5)
              to (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, dashed, in=-90, out=90]
              (1, 0)
              to (-1, 1)
              to (1, 2)
              to[in=0] (0, 2.5)
              (0, 2)
              to[out=180] (-1, 2.5)
              to (1, 3.5)
              to (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, in=-90, out=90]
              (1, 0)
              to[in=-45] (0, 3)
              (0, 1.5)
              to[out=135] (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
        =
        \vcenter{ \hbox{ \begin{tikzpicture}[x=0.75cm]
          \begin{knot}[clip width=8]
            \strand[object]
              (0, 0)
              to (0, 4.5);
            \strand[object, dashed, in=-90, out=90]
              (1, 0)
              to[in=-45] (0, 1.5)
              (0, 3)
              to[out=135] (-1, 4.5);
          \end{knot}
        \end{tikzpicture} } }
      \]
      We introduce $\sigma$ with it's inverse in step one.
      We apply naturality of $\sigma^{-1}$ in step two and one hexagon equation in step three.
      The rest follows from the correctness of the graphical calculus for general monoidal categories and cancellation of inverses.
    \end{enumerate}
  \item
    \begin{enumerate}
    \item
      \[ \begin{tikzcd}[column sep=0.5cm]
        & B
          \arrow[ld, "\rho_B^{-1}"']
          \arrow[rd, "\lambda_B^{-1}"] \\
        %
        B \otimes I
          \arrow[rr, "\sigma_{B, I}"]
          \arrow[d, "1_B \otimes f"] &&
        I \otimes B
          \arrow[d, "f \otimes 1_B"] \\
        %
        B \otimes (A \otimes C)
          \arrow[rr, "\sigma_{B, A \otimes C}"]
          \arrow[d, "\alpha_{B, A, C}^{-1}"] &&
        (A \otimes C) \otimes B
          \arrow[d, "\alpha_{A, C, B}"] \\
        %
        (B \otimes A) \otimes C
          \arrow[d, "\sigma_{B, A} \otimes 1_C"] &&
        A \otimes (C \otimes B)
          \arrow[d, "1_A \otimes \sigma_{C, B}"] \\
        %
        (A \otimes B) \otimes C
          \arrow[rr, "\alpha_{A, B, C}"] &&
        A \otimes (B \otimes C)
      \end{tikzcd} \]
      The left outer path describes the first diagram and the right outer path describes the second diagram.
      From top to bottom, the parts commute by the equality from (iii), naturality of $\sigma$ and one hexagon equality.

    \item
      \[ \begin{tikzcd}[column sep=1.2cm]
        (A \otimes B) \otimes C
          \arrow[r, "\alpha_{A, B, C}"]
          \arrow[d, "\sigma_{A, B} \otimes 1_C"] &
        A \otimes (B \otimes C)
          \arrow[d, "1_A \otimes \sigma_{B, C}"] \\
        %
        (B \otimes A) \otimes C
          \arrow[d, "\alpha_{B, A, C}"] &
        A \otimes (C \otimes B)
          \arrow[d, "\alpha_{A, C, B}^{-1}"] \\
        %
        B \otimes (A \otimes C)
          \arrow[r, "\sigma_{B, A \otimes C}"]
          \arrow[d, "1_B \otimes \sigma_{A, C}"] &
        (A \otimes C) \otimes B
          \arrow[d, "\sigma_{A, C} \otimes 1_B"] \\
        %
        B \otimes (C \otimes A)
          \arrow[r, "\sigma_{B, C \otimes A}"]
          \arrow[d, "\alpha_{B, C, A}^{-1}"] &
        (C \otimes A) \otimes B
          \arrow[d, "\alpha_{C, A, B}"] \\
        %
        (B \otimes C) \otimes A
          \arrow[d, "\sigma_{B, C} \otimes 1_A"] &
        C \otimes (A \otimes B)
          \arrow[d, "1_C \otimes \sigma_{A, B}"] \\
        %
        (C \otimes B) \otimes A
          \arrow[r, "\alpha_{C, B, A}"] &
        C \otimes (B \otimes A)
      \end{tikzcd} \]
      The left outer path describes the first diagram and the right outer path describes the second diagram.
      The parts of the diagram commute by the hexagon equalities and naturality of $\sigma$.

    \item
      \[ \begin{tikzcd}
        I \otimes A
          \arrow[rrrrrr, bend left, "\sigma"]
          \arrow[r, "\sigma"]
          \arrow[rrdd, "\rho^{-1}"']
          \arrow[rrrddd, bend right, "\lambda"'] &
        A \otimes I
          \arrow[rrrr, bend left, "\sigma"]
          \arrow[r, "1 \otimes \lambda^{-1}"]
          \arrow[rd, "\rho^{-1}"'] &
        A \otimes (I \otimes I)
          \arrow[rr, "\sigma"]
          \arrow[d, "\alpha^{-1}"] &&
        (I \otimes I) \otimes A
          \arrow[d, "\alpha"'] &
        I \otimes A
          \arrow[l, "\lambda^{-1} \otimes 1"']
          \arrow[ld, "\lambda^{-1}"] &
        A \otimes I
          \arrow[l, "\sigma"']
          \arrow[lldd, "\lambda^{-1}"]
          \arrow[lllddd, bend left, "\rho"] \\
        %
        && (A \otimes I) \otimes I
          \arrow[d, "\sigma \otimes 1"] &&
        I \otimes (I \otimes A)
          \arrow[d, "1 \otimes \sigma"'] \\
        %
        && (I \otimes A) \otimes I
          \arrow[rr, "\alpha"] &&
        I \otimes (A \otimes I) \\
        %
        &&& A
      \end{tikzcd} \]
      The rectangle in the middle represents one hexagon equality.
      The section above it commutes by naturality of $\sigma$.
      The squares to it's right and left commute by naturality of $\lambda$ and $\rho$.
      Everything else commutes trivially or by the coherence theorem for general monoidal categories.
    \end{enumerate}
  \end{enumerate}
\end{answer}

\begin{exercise}
  Show that $\Set$ is a symmetric monoidal category under $I \coloneqq \emptyset$ and $A \otimes B \coloneqq A + B + (A \times B)$, where we write $\times$ for Cartesian product of sets and $+$ for disjoint union of sets.
\end{exercise}

\begin{answer}
  You can define the associator, unitors and symmetry by composition of the corresponding morphisms for products and coproducts as well as distributivity.
  As compositions of natural isomorphisms they will also be natural isomorphisms.
  You can then check the triangle, pentagon and hexagon equalities and show that the symmetry is self-inverse (Application of a proof assistant adviced).

  Another approach is to look at a category that includes $\Set$ as subcategory but makes the monoid structure more obvious.
  We observe that $I$ and $\otimes$ coincide with the terminal object and the product functor from the category of sets with partial functions.
  Liker every terminal object and product, they give rise to a monoidal structure.
  Since the associator, unitors and symmetry are isomorphisms, they are total and contained in $\Set$.
  Thus the monoidal structure exists also $\Set$.
\end{answer}

\begin{exercise}
  Assume that we have morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon, \varepsilon'} I$ in a monoidal category such that $\eta$ and $\varepsilon$ satisfy one snake equation and $\eta$ and $\varepsilon'$ satisfy the other snake equation.
  Show that this already implies $\varepsilon = \varepsilon'$.
\end{exercise}

\begin{answer}
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (epsilon_left) at (0, 1) {};
      \node[phantom morphism] (epsilon_right) at (1, 1) {};
      \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
      \node at (epsilon) {$\varepsilon$};

      \coordinate (in1) at (0, -0.5);
      \coordinate (in2) at (1, -0.5);
      \draw[object] (in1) to (epsilon_left);
      \draw[object] (in2) to (epsilon_right);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (epsilon_left) at (0, 1) {};
      \node[phantom morphism] (epsilon_right) at (1, 1) {};
      \node[morphism, fit=(epsilon_left) (epsilon_right), inner sep=0] (epsilon) {};
      \node at (epsilon) {$\varepsilon$};

      \node[phantom morphism] (eta_left) at (1, 0) {};
      \node[phantom morphism] (eta_right) at (2, 0) {};
      \node[morphism, fit=(eta_left) (eta_right), inner sep=0] (eta) {};
      \node at (eta) {$\eta$};

      \node[phantom morphism] (epsilon'_left) at (2, 1) {};
      \node[phantom morphism] (epsilon'_right) at (3, 1) {};
      \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
      \node at (epsilon') {$\varepsilon'$};

      \coordinate (in1) at (0, -0.5);
      \coordinate (in2) at (3, -0.5);
      \draw[object] (in1) to (epsilon_left);
      \draw[object] (eta_left) to (epsilon_right);
      \draw[object] (eta_right) to (epsilon'_left);
      \draw[object] (in2) to (epsilon'_right);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (epsilon'_left) at (2, 1) {};
      \node[phantom morphism] (epsilon'_right) at (3, 1) {};
      \node[morphism, fit=(epsilon'_left) (epsilon'_right), inner sep=0] (epsilon') {};
      \node at (epsilon') {$\varepsilon'$};

      \coordinate (in1) at (2, -0.5);
      \coordinate (in2) at (3, -0.5);
      \draw[object] (in1) to (epsilon'_left);
      \draw[object] (in2) to (epsilon'_right);
    \end{tikzpicture} } }
  \]
  In the first step we apply the snake equation for $\varepsilon$ and in the second step the snake equation for $\varepsilon'$.
\end{answer}

\begin{exercise}
  In a monoidal category, show that
  \begin{enumerate}
    \item if an initial object $0$ exists and $L \dashv R$, then $L \otimes 0 \iso 0 \iso 0 \otimes R$;
    \item if a terminal object $1$ exists and $L \dashv R$, then $R \otimes 1 \iso 1 \iso 1 \otimes L$.
  \end{enumerate}
\end{exercise}

\begin{answer}
  We only show the equivalence $L \otimes 0 \iso 0$.
  The other equivalences are analogous.

  We define functions in both directions by
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (1, 1) {};
    \node[phantom morphism] (f_right) at (2, 1) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$!$};

    \node[phantom morphism] (L_mid) at (0, 1) {};

    \coordinate (in_L) at (0, 0);
    \coordinate (in_0) at (1.5, 0);
    \coordinate (out) at (2, 2);

    \draw[object] (in_L)
      to node[auto] {$L$} (L_mid.south)
      to (L_mid.north)
      to[cap object] (f_left);
    \path (f_left) to node[auto, swap] {$R$} (f_left |- out);
    \draw[object] (in_0) to node[auto] {$0$} (f);
    \draw[object] (f_right) to node[auto, swap] {$0$} (out);
  \end{tikzpicture} \]
  and
  \[ \begin{tikzpicture}
    \node[phantom morphism] (f_left) at (0, 0) {};
    \node[phantom morphism] (f_right) at (1, 0) {};
    \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
    \node at (f) {$!$};

    \draw[object] (0.5, -1) to node[auto] {$0$} (f);
    \draw[object] (f_left) to node[auto] {$L$}  (0, 1);
    \draw[object] (f_right) to node[auto] {$0$}  (1, 1);
  \end{tikzpicture}. \]
  In one direction they are trivially inverse because morphisms to $0$ are unique.
  In the other direction it follows by
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$!$};

      \node[phantom morphism] (g_left) at (1.5, 2) {};
      \node[phantom morphism] (g_right) at (2.5, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$!$};

      \node[phantom morphism] (L_mid) at (0, 1) {};

      \coordinate (in_L) at (0, 0);
      \coordinate (in_0) at (1.5, 0);
      \coordinate (out_L) at (1.5, 3);
      \coordinate (out_0) at (2.5, 3);

      \draw[object] (in_L)
        to node[auto] {$L$} (L_mid.south)
        to (L_mid.north)
        to[cap object] (f_left);
      \path (f_left) to node[auto, swap] {$R$} (f_left |- g.south);
      \draw[object] (in_0) to node[auto] {$0$} (f);
      \draw[object] (f_right) to node[auto, swap] {$0$} (g);
      \draw[object] (g_left) to node[auto] {$L$} (out_L);
      \draw[object] (g_right) to node[auto] {$0$} (out_0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2.5, 1) {};
      \node[morphism, fit=(f_left) (f_right), inner sep=0] (f) {};
      \node at (f) {$!$};

      \node[phantom morphism] (g_left) at (2, 2) {};
      \node[phantom morphism] (g_right) at (3, 2) {};
      \node[morphism, fit=(g_left) (g_right), inner sep=0] (g) {};
      \node at (g) {$!$};

      \node[phantom morphism] (L_mid_1) at (0, 1) {};
      \node[phantom morphism] (L_mid_2) at (0, 2) {};
      \node[phantom morphism] (R_mid) at (1, 2) {};

      \coordinate (in_L) at (0, 0);
      \coordinate (in_0) at (1.75, 0);
      \coordinate (out_L) at (2, 3);
      \coordinate (out_0) at (3, 3);

      \draw[object] (in_L)
        to node[auto] {$L$} (L_mid_1.south)
        to (L_mid_2.north)
        to[cap object] (R_mid)
        to (f_left);
      \path (R_mid) to node[auto, swap] {$R$} (R_mid |- out_L);
      \draw[object] (in_0) to node[auto] {$0$} (f);
      \draw[object] (f_right) to node[auto] {$0$} (g);
      \draw[object] (g_left) to node[auto, swap] {$L$} (out_L);
      \draw[object] (g_right) to node[auto, swap] {$0$} (out_0);

      \node[highlight box, fit=(f) (g)] {};
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (f_left) at (1, 1) {};
      \node[phantom morphism] (f_right) at (2.5, 1) {};
      \node[phantom morphism, fit=(f_left) (f_right), inner sep=0] (f) {};

      \node[phantom morphism] (g_left) at (2, 2) {};
      \node[phantom morphism] (g_right) at (3, 2) {};
      \node[phantom morphism, fit=(g_left) (g_right), inner sep=0] (g) {};

      \node[phantom morphism] (L_mid_1) at (0, 1) {};
      \node[phantom morphism] (L_mid_2) at (0, 2) {};
      \node[phantom morphism] (R_mid) at (1, 2) {};
      \node[phantom morphism] (L_mid_3) at (2, 2) {};

      \coordinate (in_L) at (0, 0);
      \coordinate (in_0) at (3, 0);
      \coordinate (out_L) at (2, 3);
      \coordinate (out_0) at (3, 3);

      \draw[object] (in_L)
        to node[auto] {$L$} (L_mid_1.south)
        to (L_mid_2.south)
        to (L_mid_2.north)
        to[cap object] (R_mid)
        to (R_mid.south)
        to[cup object] (L_mid_3)
        to (L_mid_3.north)
        to node[auto, swap] {$L$} (out_L);
      \path (R_mid) to node[auto, swap] {$R$} (R_mid |- out_L);
      \draw[object] (in_0)
        to node[auto] {$0$} (in_0 |- f.south)
        to (out_0 |- g.south)
        to (out_0 |- g.north)
        to node[auto, swap] {$0$} (out_0);

      \node[highlight box, fit=(f_left) (g_right)] {};
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (L_mid) at (0, 1) {};
      \node[phantom morphism] (0_mid) at (1, 1) {};

      \draw[object] (0, 0)
        to node[auto] {$L$} (L_mid.south)
        to (0, 3);
      \draw[object] (1, 0)
        to node[auto] {$0$} (0_mid.south)
        to (1, 3);
    \end{tikzpicture} } }.
  \]
  In the second step we use again that morphisms to $0$ are unique.
\end{answer}

\begin{definition}{Idempotent, splitting}
  An endomorphism $A \xrightarrow{f} A$ is called \demph{idempotent} when $f \of f = f$. An idempotent $A \xrightarrow{f} A$ \demph{splits} when there exists an object $B$ and morphisms $A \xrightarrow{g} B$ and $B \xrightarrow{h} A$ such that $f = h \of g$ and $g \of h = 1$.
\end{definition}

\begin{exercise}
  Show that in a monoidal category where all idempotents split, if there are morphisms $I \xrightarrow{\eta} R \otimes L$ and $L \otimes R \xrightarrow{\varepsilon} I$ satisfying one of the snake equations, then $L$ has a right dual (not necessarily given by $R$ with $\eta$ and $\varepsilon$).
\end{exercise}

\begin{answer}
  For clarity, we draw $\eta$ as
  \[ \begin{tikzpicture}
    \draw[object] (0, 0.5)
      to node[auto, swap] {$R$} (0, 0)
      to[cup object] (1, 0)
      to node[auto] {$L$} (1, 0.5);
    \end{tikzpicture} \]
    and $\varepsilon$ as
    \[ \begin{tikzpicture}
      \draw[object] (0, -0.5)
        to node[auto] {$L$} (0, 0)
        to[cap object] (1, 0)
        to node[auto, swap] {$R$} (1, -0.5);
    \end{tikzpicture}. \]

    Assume that we have the snake equality
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (0, 2)
        to (0, 1)
        to[cup object] (1, 1)
        to[cap object] (2, 1)
        to (2, 0);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \draw (0, 2) to (0, 0);
    \end{tikzpicture} } }. \]

    Then the other snake is idempotent by
    \[ \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (0, -1)
        to (0, 0)
        to[cap object] (1, 0)
        to[cup object] (2, 0)
        to (2, 2)
        to[cap object] (3, 2)
        to[cup object] (4, 2)
        to (4, 3);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (0, -1)
        to (0, 2)
        to[cap object] (1, 2)
        to (1, 1)
        { [local bounding box=inner snake]
          node[phantom morphism] {}
          to[cup object] (2, 1)
          to[cap object] (3, 1)
          node[phantom morphism] {}
        }
        to (3, 0)
        to[cup object] (4, 0)
        to (4, 3);
      \node[highlight box, fit=(inner snake)] {};
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \draw[object] (0, -1)
        to (0, 2)
        to[cap object] (1, 2)
        to (1, 1)
        { [local bounding box=inner snake 2]
          node[phantom morphism] {}
        }
        to (1, 0)
        to[cup object] (2, 0)
        to (2, 3);
      % \node[highlight box, fit=(inner snake 2) (inner snake 2 |- inner snake.north) (inner snake 2 |- inner snake.south)] {};
    \end{tikzpicture} } }.
  \]
  Hence there are morphisms $g$ and $h$ such that
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (out) at (0, 3) {};
      \node[morphism] (g) at (0, 2) {$g$};
      \node[morphism] (h) at (0, 1) {$h$};
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \draw[object] (in) to (h) to (g) to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (out) at (0, 3) {};
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \draw[object] (in) to (out);
    \end{tikzpicture} } }
  \]
  and
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (out) at (0, 3) {};
      \node[morphism] (h) at (0, 2) {$h$};
      \node[morphism] (g) at (0, 1) {$g$};
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \draw[object] (in) to (g) to (h) to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (out) at (2, 3) {};
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \draw[object] (in)
        to (0, 1.5)
        to[cap object] (1, 1.5)
        to[cup object] (2, 1.5)
        to (out);
    \end{tikzpicture} } }.
  \]
  We use those morphisms to define a unit
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (out_left) at (0, 1) {};
      \node[phantom morphism, overlay] (out_right) at (1, 1) {};
      \node[morphism] (g) at (1, 0) {$g$};
      \draw[object] (out_left)
        to node[auto, swap] {$R$} (0, 1 |- g.north)
        to (0, 1 |- g.south)
        to[cup object] (g.south)
        (g.north) to node[auto] {$L$} (out_right);
    \end{tikzpicture} } }
  \]
  and a counit
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[morphism] (h) at (0, 1) {$h$};
      \node[phantom morphism, overlay] (in_left) at (0, 0) {};
      \node[phantom morphism, overlay] (in_right) at (1, 0) {};
      \draw[object] (in_left)
        to node[auto] {$L$} (h)
        to[cap object] (1, 0 |- h.north)
        to (1, 0 |- h.south)
        to node[auto, swap] {$R$} (in_right);
    \end{tikzpicture} } },
  \]
  that exhibit a duality $L \dashv R$.
  The snake equations hold by
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (out) at (-1, 2.5) {};
      \node[morphism] (h) at (0, 1) {$h$};
      \node[morphism] (g) at (0, 0) {$g$};
      \node[phantom morphism] (in) at (1, -1.5) {};
      \draw[object] (in)
        to (in |- h.north)
        to[cap object] (h.north)
        (h.south) to (g)
        to[cup object] (out |- g.south)
        to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (out) at (-1, 2.5) {};
      \node[phantom morphism] (h) at (2, 1) {};
      \node[phantom morphism] (g) at (0, 0) {};
      \node[phantom morphism] (in) at (3, -1.5) {};
      \draw[object] (in)
        to (in |- h.north)
        to[cap object] (h.north)
        to (2, 0.5)
        to[cup object] (1, 0.5)
        to[cap object] (0, 0.5)
        to (g.south)
        to[cup object] (out |- g.south)
        to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (out) at (0, 4) {};
      \node[phantom morphism] (in) at (2, 0) {};
      \draw[object] (in)
        to (2, 2)
        to[cap object] (1, 2)
        to[cup object] (0, 2)
        to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism] (out) at (0, 4) {};
      \node[phantom morphism] (in) at (0, 0) {};
      \draw[object] (in) to (out);
    \end{tikzpicture} } }
  \]
  and
  \[
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (in) at (0, -0.5) {};
      \node[morphism] (h) at (0, 1) {$h$};
      \node[morphism] (g) at (2, 3) {$g$};
      \node[phantom morphism, overlay] (out) at (2, 4.5) {};
      \draw[object] (in)
        to (h)
        to (0, 2)
        to[cap object] (1, 2)
        to[cup object] (2, 2)
        to (g)
        to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \node[morphism] (h1) at (0, 1) {$h$};
      \node[morphism] (g1) at (0, 2) {$g$};
      \node[morphism] (h2) at (0, 3) {$h$};
      \node[morphism] (g2) at (0, 4) {$g$};
      \node[phantom morphism, overlay] (out) at (0, 5) {};
      \draw[object] (in)
        to (h1)
        to (g1)
        to (h2)
        to (g2)
        to (out);
    \end{tikzpicture} } }
    =
    \vcenter{ \hbox{ \begin{tikzpicture}
      \node[phantom morphism, overlay] (in) at (0, 0) {};
      \node[phantom morphism, overlay] (out) at (0, 5) {};
      \draw[object] (in) to (out);
    \end{tikzpicture} } }.
  \]
\end{answer}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: